---
title: "EJA Population Genomics Low coverage"
author: "Arthur Oulès"
format: html
editor: source
---
# Load packages

```{r}
#| echo: false

c("here", "tidyverse",
  "pcadapt", "vcfR", "adegenet", "ape", "LEA",
  "ncdf4", "rnaturalearth", "rnaturalearthdata", "rnaturalearthhires", "marmap",
  "rentrez", "xml2",
  "patchwork", "plotly", "sf", "ggrepel", "ggpubr", "ggspatial", "ggtree", "RColorBrewer", "viridis",  "scales", "ggnewscale", "terra", "tidyterra") |>
  lapply(library, character.only = TRUE)

# Data analysis


# External data acquisition for maps


# External data acquisition for outliers BLAST
 #Allows xml file manipulation

# Making plots and maps

```

# pcadapt analysis

## Load data analysis functions

```{r}
source(here("functions", "tidy_functions.R"))
```

## All populations

### Convert vcf to bed file using plink2

```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

./plink2 --vcf 8-Embiotoca_filtered.vcf.gz --make-bed --out 8-Embiotoca_filtered
```

### Preanalysis to determine optimal K value

```{r}
EJA_lcwgs <- read.pcadapt(
  here("data", "lcwgs", "8-Embiotoca_filtered.bed"),
  type = "bed"
)

pcadapt(input = EJA_lcwgs, K = 20) |>
  plot(option = "screeplot")
```

K = 4 seems correct

### Analysis using K = 4

```{r}
popmap_lcwgs <- c(
  rep("bodega_bay", 5), # BB
  rep("big_creek", 13), # BIGC -> BCR
  rep("catalina_island", 11), # CAT
  rep("elkhorn", 6), # ELK
  rep("guadalupe_island", 10), # GUA
  rep("pacific_grove", 13), # HOP -> Pacific Grove PGR 
  rep("isla_san_jeronimo", 11), # ISJ
  rep("laguna_beach", 13), # LB
  rep("point_dume", 5), # PD
  rep("santa_barbara_island", 13), # SB
  rep("san_clemente_island", 10), # SCL
  rep("la_jolla_san_diego", 13), # SD
  rep("tomales_bay", 1), # TB
  rep("redondo_beach", 10), # RB
  rep("santa_cruz_harbour", 10) # SCH
)

pcadapt_lcwgs <- EJA_lcwgs |> pcadapt(K = 4)
```

### Exploratory plots

```{r}
pcadapt_lcwgs |> plot(option = "scores", pop = popmap_lcwgs, i = 1, j = 2)
pcadapt_lcwgs |> plot(option = "scores", pop = popmap_lcwgs, i = 2, j = 3)
pcadapt_lcwgs |> plot(option = "scores", pop = popmap_lcwgs, i = 3, j = 4)

pcadapt_lcwgs |> plot(option = "qqplot")

pcadapt_lcwgs |> plot(option = "manhattan")

pcadapt_lcwgs$pvalues |> hist(
  xlab   = "p-values",
  main   = NULL,
  breaks = 50,
  col    = "orange"
)
```

## Generating whitelist

```{r}
#| echo: false
Indiv_list <- here("data", "lcwgs", "EJA_lcwgs_gt.RDS") |>
  read_rds() |>
  _$Indiv

Indiv_list |>
  write_rds(file = here("data", "lcwgs", "Individual_list.rds"))
```

```{r}
Indiv_list |> 
  unique() |>
  {\(x) {
    x[
      !(grepl(pattern = "SCL", x = x) |
        grepl(pattern = "GUA", x = x) |
        grepl(pattern = "CAT", x = x))
    ]
    }}() |> 
  write_lines(here("plink2", "whitelist_pipi.txt"))
```

## Excluding islanders
### Generating bed file
```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

./plink2 --bfile 8-Embiotoca_filtered --keep whitelist.txt --make-bed --out 8-Embiotoca_filtered_whitelisted
```

### Preanalysis to determine optimal K value

```{r}
EJA_lcwgs_coast <- read.pcadapt(
  here("data", "lcwgs", "8-Embiotoca_filtered_whitelisted.bed"),
  type = "bed"
)

pcadapt(input = EJA_lcwgs_coast, K = 20) |> plot(option = "screeplot")
```

K = 4 seems correct

### Analysis using K = 3

```{r}
popmap_lcwgs_coast <- c(
  rep("bodega_bay",            5), # BB
  rep("big_creek",            13), # BIGC -> BCR
  rep("elkhorn",               6), # ELK
  rep("pacific_grove",        13), # HOP -> Pacific Grove PGR
  rep("isla_san_jeronimo",    11), # ISJ
  rep("laguna_beach",         13), # LB
  rep("point_dume",            5), # PD
  rep("santa_barbara_island", 13), # SB
  rep("la_jolla_san_diego",   13), # SD
  rep("tomales_bay",           1), # TB
  rep("redondo_beach",        10), # RB
  rep("santa_cruz_harbour",   10)  # SCH
)

pcadapt_lcwgs_coast <- EJA_lcwgs_coast |> pcadapt(K = 3)
```

### Associated exploratory plots

```{r}
coast_PCA_1_2 <- pcadapt_lcwgs_coast |>
  plot(option = "scores", pop = popmap_lcwgs_coast, i = 1, j = 2) +
  theme_minimal()

save_open_plot(
  here("output", "plots", "lcwgs", "coast_PCA_1_2.pdf"),
  plot   = coast_PCA_1_2,
  width  = 10,
  height = 8
)

coast_PCA_2_3 <- pcadapt_lcwgs_coast |>
  plot(option = "scores", pop = popmap_lcwgs_coast, i = 2, j = 3) +
  theme_minimal()

save_open_plot(
  here("output", "plots", "lcwgs", "coast_PCA_2_3.pdf"),
  plot   = coast_PCA_2_3,
  width  = 10,
  height = 8
)

pcadapt_lcwgs_coast |> plot(option = "qqplot") + theme_minimal()

coast_manhattan <- pcadapt_lcwgs_coast |>
  plot(option = "manhattan") +
  theme_minimal()

save_open_plot(
  here("output", "plots", "lcwgs", "coast_manhattan.pdf"),
  plot   = coast_manhattan,
  width  = 16,
  height = 4
)

pcadapt_lcwgs_coast$pvalues |> hist(
  xlab   = "p-values",
  main   = NULL,
  breaks = 50,
  col    = "orange"
)
```

```{r}
rm(EJA_lcwgs_coast)
```


## Using only islanders

### Generating bed file
```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

./plink2 --bfile 8-Embiotoca_filtered --remove whitelist.txt --make-bed --out 8-Embiotoca_filtered_blacklisted
```

### Preanalysis to determine optimal K value

```{r}
EJA_lcwgs_islands <- read.pcadapt(
  here("data", "lcwgs", "8-Embiotoca_filtered_blacklisted.bed"),
  type = "bed"
)

pcadapt(input = EJA_lcwgs_islands, K = 2) |> plot(option = "screeplot")
```

### Analysis using K = 2

```{r}
popmap_lcwgs_islands <- c(
  rep("catalina_island", 11), # CAT
  rep("guadalupe_island", 10), # GUA
  rep("san_clemente_island", 10) # SCL
)

pcadapt_lcwgs_islands <- EJA_lcwgs_islands |> pcadapt(K = 2)
```

### Associated exploratory plots

```{r}
islands_PCA_1_2 <- pcadapt_lcwgs_islands |>
  plot(option = "scores", pop = popmap_lcwgs_islands, i = 1, j = 2) +
  theme_minimal()

save_open_plot(
  here("output", "plots", "lcwgs", "islands_PCA_1_2.pdf"),
  plot   = islands_PCA_1_2,
  width  = 10,
  height = 8
)

pcadapt_lcwgs_islands |>
  plot(option = "qqplot") +
  theme_minimal()

islands_manhattan <- pcadapt_lcwgs_islands |>
  plot(option = "manhattan") +
  theme_minimal()

save_open_plot(
  here("output", "plots", "lcwgs", "islands_manhattan.pdf"),
  plot   = islands_manhattan,
  width  = 16,
  height = 4
)

pcadapt_lcwgs_islands$pvalues |> hist(
  xlab   = "p-values",
  main   = NULL,
  breaks = 50,
  col    = "orange"
)
```

# Admixture analysis

## Convert data to ped format for LEA import using plink2
```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2
./plink2 --vcf 8-Embiotoca_filtered.vcf.gz --export ped --out 8-Embiotoca_filtered
```

## Read and convert data
```{r}
#| eval: false
# EJA_lcwgs_vcf <- here("data", "lcwgs", "8-Embiotoca_filtered.vcf.gz") |>
#   read.vcfR()
# EJA_lcwgs_vcf |> write_rds(here("data", "lcwgs", "EJA_lcwgs_vcfR.RDS"))

EJA_lcwgs_vcf <- read_rds(here("data", "lcwgs", "EJA_lcwgs_vcfR.RDS"))

# EJA_lcwgs_gt <- EJA_lcwgs_vcf |> extract_gt_tidy()
# EJA_lcwgs_gt |> write_rds(here("data", "lcwgs", "EJA_lcwgs_gt.RDS"))

EJA_lcwgs_gt <- read_rds(here("data", "lcwgs", "EJA_lcwgs_gt.RDS"))

ped2geno(
  input.file  = here("plink2", "8-Embiotoca_filtered.ped"),
  output.file = here("data", "lcwgs", "8-Embiotoca_filtered.geno")
)
```

## Run analysis
```{r}
snmf_analysis <- snmf(
  here("data", "lcwgs", "8-Embiotoca_filtered.geno"),
  project = "continue",
  K       = 5:8,
  entropy = TRUE
)
```

## Plot cross entropy to chose value of K
```{r}
plot(snmf_analysis)
```
K = 6 has the minimum cross-entropy.

## Rerun multiple times with K = 6
```{r}
snmf_analysis <- snmf(
  here("data", "lcwgs", "8-Embiotoca_filtered.geno"),
  project     = "continue",
  repetitions = 100,
  K           = 6,
  entropy     = TRUE
)
```

## Plot results for K = 6
```{r}
barchart(
  snmf_analysis,
  K         = 6,
  run       = which.min(cross.entropy(snmf_analysis, K = 6)),
  col       = viridis(n = 6, option = "turbo"),
  sort.by.Q = FALSE
)
```

# Figures

## Load plotting functions

```{r}
source(here("functions", "plot_functions.R"))
```

## Figure 1 - Manhattan plot home-made
### Prepare all data
```{r}
chromosome_names <- EJA_lcwgs_vcf |>
  getFIX() |>
  as_tibble() |>
  select(CHROM, POS) |>
  mutate(
    POS  = as.numeric(POS),
    rank = row_number()
  )

labs_pos <- chromosome_names |>
  mutate(labs_pos = row_number()) |>
  distinct(CHROM, .keep_all = TRUE) |>
  pull(var = labs_pos)

chromosome_names <- chromosome_names |>
  mutate(
    pvalues_all    = Get_pvalues(pcadapt_lcwgs),
    pvalues_coast  = Get_pvalues(pcadapt_lcwgs_coast),
    pvalues_island = Get_pvalues(pcadapt_lcwgs_islands),
    lab_position   = if_else(row_number() %in% labs_pos, TRUE, FALSE),
    coloured       = CHROM |>
      {\(x) factor(x, levels = unique(x))}() |>
      as.integer() |>
      (`%%`)(2)
  )

rm(labs_pos)
```

### Plot every populations
```{r}
manhattan_plot_lcwgs <- chromosome_names |>
  manhattan_plot_custom_2(pvalues = pvalues_all)

ggsave(
  here("output", "plots", "lcwgs", "manhattan_plot_custom_lcwgs.png"),
  plot   = manhattan_plot_lcwgs,
  width  = 6000,
  height = 2000,
  units = "px"
)
```

### Plot coastal populations
```{r}
manhattan_plot_lcwgs_coast <- chromosome_names |>
  manhattan_plot_custom_2(pvalues = pvalues_coast)

ggsave(
  here("output", "plots", "lcwgs", "manhattan_plot_custom_lcwgs_coast.png"),
  plot   = manhattan_plot_lcwgs_coast,
  width  = 6000,
  height = 2000,
  units = "px"
)
```

### Plot island populations
```{r}
manhattan_plot_lcwgs_islands <- chromosome_names |>
  manhattan_plot_custom_2(pvalues = pvalues_island)

ggsave(
  here("output", "plots", "lcwgs", "manhattan_plot_custom_lcwgs_islands.png"),
  plot   = manhattan_plot_lcwgs_islands,
  width  = 6000,
  height = 2000,
  units = "px"
)
```

### Clean memory
```{r}
rm(EJA_lcwgs_vcf, chromosome_names,
   manhattan_plot_lcwgs, manhattan_plot_lcwgs_coast, manhattan_plot_lcwgs_islands)
```

## Figure 2 - PCA & map
### PCA plot

```{r}
PCA <- PCA_plot(pcadapt_output = pcadapt_lcwgs, popmap = popmap_lcwgs)
PCA

save_open_plot(
  path   = here("output", "plots", "lcwgs", "PCA.pdf"),
  plot   = PCA,
  width  = 9,
  height = 9
)

PCA_plot(
  pcadapt_output = pcadapt_lcwgs,
  popmap         = popmap_lcwgs,
  x_offsets      = rep(.01, 15),
  y_offsets      = rep(.05, 15)
) |>
  ggplotly()

PCA
```

```{r}
PCA_crop <- PCA + xlim(-.065, -.05) + ylim(.1125, .1275) +
  coord_fixed(ratio = 1) +
  theme(
    legend.position = "none",
    axis.title      = element_blank(),
    plot.background = element_rect(fill = "white", linetype = "longdash")
  )
PCA_crop$layers[[2]] <- NULL

PCA_and_crop <- PCA + annotate(
  "rect",
  xmin  = -.07, xmax = -.04,
  ymin  = .1075, ymax = .1325,
  alpha = 0, colour = "black", linetype = "longdash"
) + 
theme(legend.position = "none") +
inset_element(PCA_crop, left = 0.6, bottom = 0.6, right = 1, top = 1)

save_open_plot(
  path   = here("output", "plots", "lcwgs", "pca_crop.pdf"),
  plot   = PCA_and_crop,
  width  = 9,
  height = 9
)
```

### Map - SST and current 
#### Formatting Copernicus data
```{r}
#| echo: false

currents_tibble <- left_join(
  rast(
    x     = here("data", "Copernicus", "merged_uv",
                 "cmems_mod_glo_phy-cur_anfc_0.083deg_P1D-m_1721914334862.nc"),
    subds = "uo"
  ) |>
    as.data.frame(xy = TRUE, time = TRUE, na.rm = FALSE) |> 
    pivot_longer(cols = -c(x, y), names_to = "Date", values_to = "u") |> 
    mutate(Date = as.Date(Date)),
  rast(
    x    = here("data", "Copernicus", "merged_uv",
                "cmems_mod_glo_phy-cur_anfc_0.083deg_P1D-m_1721914334862.nc")
  ) |>
    as.data.frame(xy = TRUE, time = TRUE, na.rm = FALSE) |> 
    pivot_longer(cols = -c(x, y), names_to = "Date", values_to = "v") |> 
    mutate(Date = as.Date(Date))
)

currents_tibble_annual_mean <- currents_tibble |> 
  group_by(x, y) |> 
  summarise(u_mean = mean(u), v_mean = mean(v)) |> 
  ungroup()

currents_tibble_annual_mean_lres <- left_join(
  rast(
    x     = here("data", "Copernicus", "merged_uv",
                 "cmems_mod_glo_phy-cur_anfc_0.083deg_P1D-m_1721914334862.nc"),
    subds = "uo"
  ) |>
    mean() |> 
    aggregate(fact = 2, fun = "mean") |> 
    as.data.frame(xy = TRUE, na.rm = FALSE) |> 
    as_tibble() |> 
    rename(u_mean = mean),
  rast(
    x     = here("data", "Copernicus", "merged_uv",
                "cmems_mod_glo_phy-cur_anfc_0.083deg_P1D-m_1721914334862.nc"),
    subds = "vo"
  ) |>
    mean() |> 
    aggregate(fact = 2, fun = "mean") |> 
    as.data.frame(xy = TRUE, na.rm = FALSE) |> 
    as_tibble() |> 
    rename(v_mean = mean),
)

currents_tibble_annual_mean_lres |>
  write_rds(file = here("data", "Copernicus", "annual_average_currents_tb.RDS"))

rm(currents_tibble, currents_tibble_annual_mean)
```

```{r}
SST_maps <- here("data", "Copernicus", "SST_temperature_UKMO-L4_GHRSST") |> 
  list.files(full.names = TRUE) |>
  map(
    \(x) x |> 
      rast(lyrs = "analysed_sst", win = ext(-124,-113, 28.5, 38.5))
  ) |> 
  rast()
time(SST_maps) <- seq(as.Date("2023-01-01"), as.Date("2023-12-31"), by = "days")

SST_annual_mean <- SST_maps |> mean() |> (`-`)(273.15)

SST_annual_mean |>
  writeRaster(file = here("data", "Copernicus", "annual_average_SST.tif"))

rm(SST_maps)
```

```{r}
currents_tibble_annual_mean_lres <- read_rds(
  file = here("data", "Copernicus", "annual_average_currents_tb.RDS")
)

SST_annual_mean <- read_rds(
  file = here("data", "Copernicus", "annual_average_SST.RDS")
)
```

#### Landmass
```{r}
# Setting up landmass map from naturalearth
states <- ne_states(
  country     = c("united states of america", "mexico"),
  returnclass = "sf"
)

# Sampling site coordinates and misc info on map
sample_sites <- read.csv2(
    here("data","low_coverage_sampling.csv"),
    stringsAsFactors = FALSE
  ) |>
  select(Longitude, Latitude, Sampling_sites) |> 
  mutate( # Add offsets
    Sampling_sites = gsub("\\n", "\n", Sampling_sites, fixed = TRUE),
    lab_lon = Longitude + .15 + rep(0, 5),
    lab_lat = Latitude + .25 + rep(0, 5)
  )

states_labels <- data.frame(
  "long" = c(-117.5, -115.5, -113.7, -115.2),
  "lat"  = c(35.5, 36.6, 35, 30.75),
  "lab"  = c("CA", "NV", "AZ", "BC")
)
city_labels <- data.frame(
  "lat"      = c(37.773972, 34.448113),
  "long"     = c(-122.431297, -120.471439),
  "lat_lab"  = c(37.7, 34.8),
  "long_lab" = c(-122.28, -120.535),
  "lab"      = c("San Francisco", "Point\nConception")
)
ocean_labels <- data.frame(
  "long" = c(-121,-113.85),
  "lat"  = c(31,30.7),
  "lab"  = c("Pacific Ocean","Gulf\nof\nCalifornia")
)
```

#### Plot map
[Tutorial for surface temperature and current speed map](https://theoceancode.netlify.app/post/mapping_with_ggplot2/)

```{r}
lcwgs_map <- ggplot() +
  # Temperature map
  geom_spatraster(data = SST_annual_mean, aes(fill = mean)) +
  # Temperature scale
  scale_fill_viridis(
    name = "Temp.\n (°C)",
    breaks = c(12.5, 15, 17.5, 20, 22.5),
    option = "turbo"
  ) +
  # Current vectors
  geom_segment(
    data     = currents_tibble_annual_mean_lres,
    mapping  = aes(
      x    = x,
      y    = y,
      xend = x + 2*u_mean,
      yend = y + 2*v_mean
    ),
    size     = .5,
    colour   = "#4d4d4d",
    arrow    = arrow(angle = 20, length = unit(0.035, "inches"), type = "closed"),
    linejoin = "mitre",
    na.rm    = TRUE
  ) +
  # Landmass
  geom_sf(data = states) +
  # Crop map
  coord_sf(xlim = c(-124, -113), ylim = c(28.5, 38.5), expand = FALSE) +
  # Sampling sites points and labels
  new_scale("fill") +
  geom_point(
    sample_sites,
    mapping     = aes(x    = Longitude,
                      y    = Latitude,
                      fill = Sampling_sites,
                      size = 1.3),
    shape       = 21,
    show.legend = FALSE
  ) +
  geom_label_repel(
    sample_sites,
    mapping = aes(x = Longitude, y = Latitude, label = Sampling_sites)
  ) +
  # Extra cities points and labels
  geom_point(
    city_labels,
    mapping     = aes(x = long, y = lat),
    size        = 4,
    shape       = 20,
    show.legend = FALSE
  ) +
  geom_text(
    city_labels,
    mapping    = aes(x = long_lab, y = lat_lab, label = lab),
    lineheight = .7,
    hjust      = 0
  ) +
  # Ocean and state names
  geom_text(
    states_labels,
    mapping = aes(x = long, y = lat, label = lab),
    size = 5.5
  ) +
  geom_text(
    ocean_labels[2, ],
    mapping = aes(x = long, y = lat, label = lab),
    colour  = "ivory",
    size    = 5
  ) +
  geom_label(
    ocean_labels[1, ],
    mapping       = aes(x = long, y = lat, label = lab),
    colour        = "black",
    size          = 5,
    label.padding = unit(0.5, "lines")
  ) +
  # Vector scale
  geom_label(
    aes(x = -122, y = 29.5, label = ".25 m/s\n"),
    size          = 4.5,
    label.padding = unit(0.5, "lines")
  ) +
  geom_segment(
    aes(x = -122.25, y = 29.25, xend = -121.75, yend = 29.25),
    size     = .5,
    arrow = arrow(angle = 20, length = unit(0.035, "inches"), type = "closed"),
    linejoin = "mitre"
  ) +
  # Cartographic scale
  geom_label(aes(x = -119.1, y = 28.71, label = "        ")) +
  annotation_scale(location = "bl", width_hint = 0.4) +
  # Orientation
  annotation_north_arrow(
    location = "tr",
    style = north_arrow_fancy_orienteering()
  ) +
  # Axes
  labs(x = "Longitude", y = "Latitude") +
  scale_x_continuous(
    breaks = c(-120, -115),
    labels = c("120°W", "115°W")
  ) +
  scale_y_continuous(
    breaks = c(30, 35),
    labels = unit_format(suffix = "°N", sep = "")
  )

save_open_plot(
  here("output", "plots", "lcwgs", "temp_current_map_lres.pdf"),
  plot   = lcwgs_map,
  width  = 9,
  height = 9
)
```

### Memory cleanup
```{r}
rm(currents_tibble_annual_mean_lres, SST_annual_mean,
   map, states,
   sample_sites, states_labels, city_labels, ocean_labels,
   current_uv_scalar)
gc()
```

### Fuse PCA plot and map
```{r}
PCA_map <- PCA_and_crop + lcwgs_map +
  plot_annotation(tag_levels = list(c("A", "", "B"))) &
  theme(
    text      = element_text(size = 14),
    axis.text = element_text(size = 12),
    plot.tag  = element_text(face = "bold")
    )

save_open_plot(
  here("output", "plots", "lcwgs", "pca_map.png"),
  plot   = PCA_map,
  width  = 17.5,
  height = 11.25
)
```

### 3D PCA
```{r}
TD_PCA_plot <- pcadapt_lcwgs$scores |>
  as_tibble() |>
  rename(PC_1 = V1, PC_2 = V2, PC_3 = V3, PC_4 = V4) |> 
  mutate(pop = popmap_lcwgs) |> 
  as.data.frame() |> 
  plot_ly(
    x = ~PC_1, y = ~PC_2, z = ~PC_3,
    color = ~pop, colors = viridis(144, option = "turbo")
  ) |>
  add_markers(size = 12)

setwd(here("output", "plots", "lcwgs"))
as_widget(TD_PCA_plot) |> saveWidget(
  file = here("output", "plots", "lcwgs", "3D_PCA.html"),
  selfcontained = TRUE)
setwd(here())

rm(TD_PCA_plot)
```

## Figure 5 - Admixture plot
Change individuals order in the file `Individuals_plot_order.txt`
```{r}
admixture_plot <- read_table(
  here("data", "lcwgs", "8-Embiotoca_filtered.snmf", "K6", "run65",
       "8-Embiotoca_filtered_r65.6.Q"),
  col_names = paste0("Pop ", seq(1, 6))
) |> 
  mutate(Individuals = Indiv_list |> unique()) |> 
  relocate(Individuals) |> 
  pivot_longer(
    -c(Individuals),
    names_to  = "Populations",
    values_to = "Ancestry proportions"
  ) |> 
  ggplot(
    aes(
      x    = factor(
        Individuals,
        levels = here("data", "lcwgs", "Individuals_plot_order.txt") |> 
          read.table() |>
          unlist() |>
          as.vector() 
      ),
      y    = `Ancestry proportions`,
      fill = Populations
    )
  ) +
  geom_col(position = "stack", width = 1) +
  scale_fill_viridis(discrete = TRUE, option = "turbo") +
  labs(x = "Individuals") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust = 1)
  )

save_open_plot(
  path   = here("output", "plots", "lcwgs", "admixture_plot_K=6.pdf"),
  plot   = admixture_plot,
  width  = 16,
  height = 6
  )

rm(Indiv_list, admixture_plot)
```

## Figure 6 - Distance tree
```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

./plink --vcf 8-Embiotoca_filtered.vcf.gz --out Embiotoca_filtered_dist_matrix --allow-extra-chr --autosome-num 95 --distance square --const-fid
```


```{r}
EJA_lcwgs_dist_IBS <- here("plink2", "Embiotoca_filtered_dist_matrix.dist") |> 
  read_table(col_names = FALSE) |> 
  as.matrix()

IDs <- here("plink2", "Embiotoca_filtered_dist_matrix.dist.id") |> 
  read_table(col_names = FALSE) |> pull(X2)

colnames(EJA_lcwgs_dist_IBS) <- IDs
rownames(EJA_lcwgs_dist_IBS) <- IDs
```

```{r}
NJ_Tree_lcwgs_IBS <- EJA_lcwgs_dist_IBS |> as.dist() |> nj()

NJ_Tree_lcwgs_IBS_annotations <- tibble(
  Indiv = NJ_Tree_lcwgs_IBS$tip.label,
  pop   = popmap_lcwgs
)

NJ_Tree_lcwgs_IBS_plot <- NJ_Tree_lcwgs_IBS |>
  ggtree(layout = "circular") %<+% NJ_Tree_lcwgs_IBS_annotations +
  geom_tiplab(aes(color = pop), size = 2.8) +
  geom_treescale(x = 800000, y = 0)

save_open_plot(
  path   = here("output", "plots", "lcwgs", "dist_tree.pdf"),
  plot   = NJ_Tree_lcwgs_IBS_plot,
  width  = 12,
  height = 9
)
```

```{r}
geo_dist <- here("data", "low_coverage_sampling.csv") |> 
  read_csv2(show_col_types = FALSE) |> 
  select(c(Code, Longitude, Latitude)) |> 
  mutate(
    pop = popmap_lcwgs |>
      unique() |>
      _[c(1, 13, 15, 4, 6, 2, 9, 14, 8, 10, 3, 11, 12, 7, 5)]
  ) |> 
  mutate(
    pop  = as.factor(pop),
    Code = as.factor(Code)
  )

geo_dist_mat <- left_join(
  x = NJ_Tree_lcwgs_IBS_annotations,
  y = geo_dist
) |> 
  select(Indiv, Longitude, Latitude) |> 
  dist() |>
  as.matrix()

colnames(geo_dist_mat) <- IDs
rownames(geo_dist_mat) <- IDs

rm(IDs)
```

```{r}
vegan::mantel(geo_dist_mat, EJA_lcwgs_dist_IBS)
plot(geo_dist_mat)
plot(EJA_lcwgs_dist_IBS)
plot(geo_dist_mat/max(geo_dist_mat), EJA_lcwgs_dist_IBS/max(EJA_lcwgs_dist_IBS))
```

```{r}
geo_df <- left_join(
    x = NJ_Tree_lcwgs_IBS_annotations,
    y = geo_dist
  ) |>
  select(Longitude, Latitude) |>
  rename(lat = Latitude, lon = Longitude) |> 
  as.data.frame()

library(vcfR)

vcf_test <- here("data", "lcwgs", "EJA_lcwgs_vcfR.RDS") |>
  readRDS() |>
  vcfR2genlight()

vcf_test@other$latlon <- geo_df
vcf_test@pop <- factor(NJ_Tree_lcwgs_IBS_annotations$pop)

vcf_test |> saveRDS(file = here("data", "lcwgs", "Embiotoca_lcwgs_gl.rds"))

library(dartR)

gl.ibd(vcf_test, coordinates = "latlon")
```