---
title: "Isolation by Distance"
author: "Arthur Oulès"
format: html
editor: source
---

## Load packages

```{r}
#| label: LoadPackages
#| echo: false

c("here", "tidyverse", "vcfR", "dartR", "adegenet",
  "topoDistance", "terra", "tidyterra", "inlabru",
  "rnaturalearth", "marmap") |>
  lapply(library, character.only = TRUE) |> invisible()
```

## Load functions

```{r}
#| label: LoadExtFunc

source(here("functions", "tidy_functions.R"))
source(here("functions", "plot_functions.R"))
```

## Helper objects

```{r}
popmap_tb <- read_csv2(here("data", "individuals_coordinates.csv")) |>
  select(
    ID     = Indiv,
    Code   = Localisation,
    Code_b = Localisation_b,
    pop    = popmap,
    pop2   = pop_map_b
  )
```

## Genetic distances

### Pairwise Hamming genetic distances

#### Compute distances in plink 1.9

```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

./plink --vcf 8-Embiotoca_filtered.vcf.gz --out Embiotoca_filtered_dist_matrix --allow-extra-chr --autosome-num 95 --distance square --const-fid
```

#### Import plink results

```{r}
EJA_lcwgs_dist <- here("plink2", "Embiotoca_filtered_dist_matrix.dist") |> 
  read_table(col_names = FALSE) |> 
  as.matrix()

IDs <- here("plink2", "Embiotoca_filtered_dist_matrix.dist.id") |> 
  read_table(col_names = FALSE) |> pull(X2)

colnames(EJA_lcwgs_dist) <- IDs
rownames(EJA_lcwgs_dist) <- IDs
```

#### Individual

##### Plot results

```{r}
ID_plot_order <- read.table(here("data", "Individuals_plot_order.txt")) |> unlist()

ID_dist_plot <- EJA_lcwgs_dist[ID_plot_order, ID_plot_order] |>
  reshape2::melt(value.name = c("dist")) |> 
  plot_matrix_notext() +
  theme(
    axis.text.x      = element_text(angle = 90, vjust = 0.5, hjust = 1),
    panel.background = element_blank()
  )

save_open_plot(
  path   = here("output", "plots", "ID_dist_plot.png"),
  plot   = ID_dist_plot,
  width  = 16,
  height = 15
)
```

##### NJ trees

```{r}
NJ_Tree_lcwgs <- EJA_lcwgs_dist |> as.dist() |> nj()

NJ_Tree_lcwgs_annotations <- tibble(
  Indiv = NJ_Tree_lcwgs$tip.label,
  pop   = popmap_lcwgs
)

NJ_Tree_lcwgs_plot <- NJ_Tree_lcwgs |>
  ggtree(layout = "circular") %<+% NJ_Tree_lcwgs_annotations +
  geom_tiplab(aes(color = pop), size = 2.8) +
  geom_treescale(x = 800000, y = 0)

save_open_plot(
  path   = here("output", "plots", "dist_tree.pdf"),
  plot   = NJ_Tree_lcwgs_plot,
  width  = 12,
  height = 9
)
```

```{r}
NJ_Tree_lcwgs_rooted <- NJ_Tree_lcwgs |> root(outgroup = "EJA_GUA20")

NJ_Tree_lcwgs_rooted_plot <- NJ_Tree_lcwgs_rooted |>
  ggtree() %<+% NJ_Tree_lcwgs_annotations +
  geom_tiplab(aes(color = pop), size = 2.8) +
  geom_treescale(x = 800000, y = 0)

save_open_plot(
  path   = here("output", "plots", "dist_tree_rooted.pdf"),
  plot   = NJ_Tree_lcwgs_rooted_plot,
  width  = 12,
  height = 12
)
```

#### All Populations

```{r}
pop_distance_matrix <- aggregate_by_population(EJA_lcwgs_dist, popmap_tb$pop)
pop_distance_matrix_sub <- Substract_diagonal(pop_distance_matrix)
```

```{r}
pop_distance_matrix[pop_order_b, pop_order_b] |>
  as.data.frame() |>
  write.csv(here("output", "EJA_all_pop_allele_distances.csv"))

pop_distance_matrix_sub[pop_order_b, pop_order_b] |>
  as.data.frame() |>
  write.csv(here("output", "EJA_all_pop_allele_distances_substracted.csv"))
```

```{r}
pop_order <- c("bodega_tomales_bay", "santa_cruz", "elkhorn", "pacific_grove", "big_creek", "santa_barbara", "point_dume", "redondo_beach", "laguna_beach", "la_jolla_san_diego", "isla_san_jeronimo", "catalina_island", "san_clemente_island", "guadalupe_island")

pop_distance_matrix_plot <- pop_distance_matrix[pop_order_b, pop_order_b] |>
  reshape2::melt(value.name = c("dist")) |> 
  plot_matrix() +
  theme(
    axis.text        = element_text(angle = 45, vjust = 0.5, hjust = 0),
    panel.background = element_blank()
  )
  
save_open_plot(
  path   = here("output", "plots", "pop_distance_matrix_plot.pdf"),
  plot   = pop_distance_matrix_plot,
  width  = 8,
  height = 7
)

pop_distance_matrix_sub_plot <- pop_distance_matrix_sub[pop_order_b, pop_order_b] |>
  reshape2::melt(value.name = c("dist")) |> 
  plot_matrix() +
  theme(
    axis.text        = element_text(angle = 45, vjust = 0.5, hjust = 0),
    panel.background = element_blank()
  )
  
save_open_plot(
  path   = here("output", "plots", "pop_distance_matrix_sub_plot.pdf"),
  plot   = pop_distance_matrix_sub_plot,
  width  = 8,
  height = 7
)
```

```{r}
NJ_Tree_lcwgs_pops_rooted <- pop_distance_matrix |> NJ() |> root("guadalupe_island")

NJ_Tree_lcwgs_pops_rooted_plot <- NJ_Tree_lcwgs_pops_rooted |>
  ggtree() +
  geom_tiplab() +
  geom_treescale(x = 800000, y = 0) +
  xlim(0, 1125000)

save_open_plot(
  path   = here("output", "plots", "dist_tree_pop_rooted.pdf"),
  plot   = NJ_Tree_lcwgs_pops_rooted_plot,
  width  = 12,
  height = 8
)

```

#### Coast + isj vs Islands

```{r}
pop_distance_matrix_2 <- aggregate_by_population(EJA_lcwgs_dist, popmap_tb$pop2)
pop_distance_matrix_2_sub <- Substract_diagonal(pop_distance_matrix_2)
```

```{r}
pop_distance_matrix_2[plot_order_coast, plot_order_coast] |>
  as.data.frame() |>
  write.csv(here("output", "EJA_coast_vs_isl_allele_distances.csv"))

pop_distance_matrix_2_sub[plot_order_coast, plot_order_coast] |>
  as.data.frame() |>
  write.csv(here("output", "EJA_coast_vs_isl_allele_distances_substracted.csv"))
```

```{r}
plot_order_coast <- c("coast_isj", "catalina_island", "san_clemente_island", "guadalupe_island")

pop_distance_matrix_2_plot <- pop_distance_matrix_2[plot_order_coast, plot_order_coast] |>
  reshape2::melt(value.name = c("dist")) |> 
  plot_matrix() +
  theme(
    axis.text        = element_text(angle = 45, vjust = 0.5, hjust = 0),
    panel.background = element_blank()
  )
  
save_open_plot(
  path   = here("output", "plots", "pop_distance_matrix_coast_vs_isl_plot.pdf"),
  plot   = pop_distance_matrix_2_plot,
  width  = 8,
  height = 7
)

pop_distance_matrix_2_sub_plot <- pop_distance_matrix_2_sub[plot_order_coast, plot_order_coast] |>
  reshape2::melt(value.name = c("dist")) |> 
  plot_matrix() +
  theme(
    axis.text        = element_text(angle = 45, vjust = 0.5, hjust = 0),
    panel.background = element_blank()
  )
  
save_open_plot(
  path   = here("output", "plots", "pop_distance_matrix_coast_vs_isl_sub_plot.pdf"),
  plot   = pop_distance_matrix_2_sub_plot,
  width  = 8,
  height = 7
)
```

### Weir and Cocherham pairwise Fst

#### Write cluster maps for plink

```{r}
read_table(here("plink2", "8-Embiotoca_filtered.fam"), col_names = FALSE) |>
  select(X1, X2) |>
  mutate(
    pop = here("data", "individuals_coordinates.csv") |>
      read_csv2() |>
      pull(Localisation),
    pop2 = here("data", "individuals_coordinates.csv") |>
      read_csv2() |>
      pull(Localisation_b)
  ) |>
  write.table(
    here("plink2", "8-Embiotoca_filtered.txt"),
    quote     = FALSE,
    row.names = FALSE,
    col.names = c("FID", "IID", "pop", "pop2"),
    sep       = "\t"
  )
```

#### All pops

##### Compute pairwise Fst with plink

```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

./plink2 --bfile 8-Embiotoca_filtered --out Embiotoca_filtered_Fst --allow-extra-chr --autosome-num 95 --const-fid --pheno 8-Embiotoca_filtered.txt --fst pop 'method='wc
```

##### Load and format results

```{r}
lookup <- popmap_tb |> select(c(Code, pop)) |> distinct(.keep_all = TRUE)

Fst_all <- here("plink2", "Embiotoca_filtered_Fst.fst.summary") |>
  read_table(skip = 1, col_names = c("Var1", "Var2", "dist")) |> 
  mutate(
    dist = as.numeric(dist)
  ) |> 
  reshape2::acast(Var1 ~ Var2)

Fst_all <- cbind(matrix(rep(NA, 14), 14, 1, dimnames = list(rownames(Fst_all), "BBA")), Fst_all)
Fst_all <- rbind(Fst_all, matrix(rep(NA, 15), 1, 15, dimnames = list("TB", colnames(Fst_all))))
Fst_all[is.na(Fst_all)] <- 0
Fst_all <- Fst_all + t(Fst_all)

colnames(Fst_all) <-  rownames(Fst_all) <- lookup$pop[match(colnames(Fst_all), lookup$Code)]
rm(lookup)
```

##### Save results

```{r}
pop_order_b <- c("bodega_bay", "tomales_bay", "santa_cruz_harbour", "elkhorn", "pacific_grove", "big_creek", "santa_barbara", "point_dume", "redondo_beach", "laguna_beach", "la_jolla_san_diego", "isla_san_jeronimo", "catalina_island", "san_clemente_island", "guadalupe_island")

Fst_all[pop_order_b, pop_order_b] |>
  as.data.frame() |>
  write.csv(here("output", "EJA_all_pop_pairwise_Fst.csv"))
```

##### Plot results

```{r}
pop_Fst_matrix_plot <- reshape2::melt(
  Fst_all[pop_order_b , pop_order_b ]*1000000,
  value.name = "dist"
) |> 
  plot_matrix() +
  scale_fill_viridis("Fst", option = "turbo") +
  theme(
    axis.text        = element_text(angle = 45, vjust = 0.5, hjust = 1),
    panel.background = element_blank()
  )

save_open_plot(
  path   = here("output", "plots", "pop_Fst_matrix_plot.pdf"),
  plot   = pop_Fst_matrix_plot,
  width  = 8,
  height = 7
)
```

#### Coast + isj vs Islands

##### Compute pairwise Fst with plink

```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

./plink2 --bfile 8-Embiotoca_filtered --out Embiotoca_filtered_coast_Fst --allow-extra-chr --autosome-num 95 --const-fid --pheno 8-Embiotoca_filtered.txt --fst pop2 'method='wc
```

##### Load and format results

```{r}
lookup <- popmap_tb |> select(c(Code_b, pop2)) |> distinct(.keep_all = TRUE)

Fst_coast_vs_isl <- here("plink2", "Embiotoca_filtered_coast_Fst.fst.summary") |>
  read_table(skip = 1, col_names = c("Var1", "Var2", "dist")) |> 
  mutate(
    dist = as.numeric(dist)
  ) |> 
  reshape2::acast(Var1 ~ Var2)

Fst_coast_vs_isl <- cbind(matrix(rep(NA, 3), 3, 1, dimnames = list(rownames(Fst_coast_vs_isl), "CAT")), Fst_coast_vs_isl)
Fst_coast_vs_isl <- rbind(Fst_coast_vs_isl, matrix(rep(NA, 4), 1, 4, dimnames = list("SCL", colnames(Fst_coast_vs_isl))))
Fst_coast_vs_isl[is.na(Fst_coast_vs_isl)] <- 0
Fst_coast_vs_isl <- Fst_coast_vs_isl + t(Fst_coast_vs_isl)

colnames(Fst_coast_vs_isl) <-  rownames(Fst_coast_vs_isl) <- lookup$pop2[match(colnames(Fst_coast_vs_isl), lookup$Code_b)]

rm(lookup)
```

##### Save results

```{r}
Fst_coast_vs_isl[plot_order_coast, plot_order_coast] |>
  as.data.frame() |>
  write.csv(here("output", "EJA_coast_vs_isl_pop_pairwise_Fst.csv"))
```

##### Plot results

```{r}
plot_order_coast <- c("coast_isj", "catalina_island", "san_clemente_island", "guadalupe_island")

pop_Fst_coast_vs_isl_matrix_plot <- reshape2::melt(Fst_coast_vs_isl[plot_order_coast, plot_order_coast]*1000000, value.name = c("dist")) |> 
  plot_matrix() +
  scale_fill_viridis("Fst", option = "turbo") +
  theme(
    axis.text        = element_text(angle = 45, vjust = 0.5, hjust = 0),
    panel.background = element_blank()
  )

save_open_plot(
  path   = here("output", "plots", "pop_Fst_coast_vs_isl_matrix_plot.pdf"),
  plot   = pop_Fst_coast_vs_isl_matrix_plot,
  width  = 8,
  height = 7
)
```

#### Combined plots

##### All pops

```{r}
pop_order_b <- c("bodega_bay", "tomales_bay", "santa_cruz_harbour", "elkhorn", "pacific_grove", "big_creek", "santa_barbara", "point_dume", "redondo_beach", "laguna_beach", "la_jolla_san_diego", "isla_san_jeronimo", "catalina_island", "san_clemente_island", "guadalupe_island")

Fst_all <- Fst_all[pop_order_b, pop_order_b]
Fst_all[lower.tri(Fst_all)] <- NA

pop_distance_matrix_sub_t <- pop_distance_matrix_sub[pop_order_b, pop_order_b]
pop_distance_matrix_sub_t[upper.tri(pop_distance_matrix_sub_t)] <- NA

df1 <- Fst_all |> reshape2::melt(value.name = "Fst") |> tibble()
df2 <- pop_distance_matrix_sub_t |> reshape2::melt(value.name = "dist") |> tibble()


all_distances_biplot <- ggplot() + 
  geom_raster(df1, mapping = aes(x = Var1, y = Var2, fill  = Fst)) +
  geom_text(df1,   mapping = aes(x = Var1, y = Var2, label = round(Fst, digits = 2))) +
  scale_fill_viridis(name = "Fst", na.value = "transparent") +
  new_scale_fill() +
  geom_raster(df2,  mapping = aes(x = Var1, y = Var2, fill  = dist/1000000)) +
  geom_text(df2,    mapping = aes(x = Var1, y = Var2, label = round(dist/1000000, digits = 2))) +
  scale_fill_viridis(name = "Distance in Mb", option = "turbo", na.value = "transparent") +
  labs(x = "", y = "") +
  scale_x_discrete(position = "top") +
  scale_y_discrete(limits = rev(levels(df1$Var2))) +
  theme(
    axis.text        = element_text(angle = 45, vjust = 0.5, hjust = 0),
    panel.background = element_blank()
  )

save_open_plot(
  path   = here("output", "plots", "pop_Fst_and_distances_all_matrix_plot.pdf"),
  plot   = all_distances_biplot,
  width  = 8,
  height = 6.5
)
```

##### Coast + isj vs islands

```{r}
Fst_coast_vs_isl_t <- Fst_coast_vs_isl[plot_order_coast, plot_order_coast]
Fst_coast_vs_isl_t[lower.tri(Fst_coast_vs_isl_t)] <- NA

pop_distance_matrix_2_sub_t <- pop_distance_matrix_2_sub[plot_order_coast, plot_order_coast]
pop_distance_matrix_2_sub_t[upper.tri(pop_distance_matrix_2_sub_t)] <- NA

df1 <- Fst_coast_vs_isl_t |> reshape2::melt(value.name = "Fst") |> tibble()
df2 <- pop_distance_matrix_2_sub_t |> reshape2::melt(value.name = "dist") |> tibble()


coast_distances_biplot <- ggplot() + 
  geom_raster(df1, mapping = aes(x = Var1, y = Var2, fill = Fst)) +
  geom_text(df1,  mapping = aes(x = Var1, y = Var2, label = round(Fst, digits = 2))) +
  scale_fill_viridis(name = "Fst", na.value = "transparent") +
  new_scale_fill() +
  geom_raster(df2,  mapping = aes(x = Var1, y = Var2, fill = dist/1000000)) +
  geom_text(df2,  mapping = aes(x = Var1, y = Var2, label = round(dist/1000000, digits = 2))) +
  scale_fill_viridis(name = "Distance in Mb", option = "turbo", na.value = "transparent") +
  labs(x = "", y = "") +
  scale_x_discrete(position = "top") +
  scale_y_discrete(limits = rev(levels(df1$Var2))) +
  theme(
    axis.text        = element_text(angle = 45, vjust = 0.5, hjust = 0),
    panel.background = element_blank()
  )

save_open_plot(
  path   = here("output", "plots", "pop_Fst_and_distances_coast_vs_isl_matrix_plot.pdf"),
  plot   = coast_distances_biplot,
  width  = 8,
  height = 6.5
)
```

### Edward distance Sendaway

See `/sendaway/Compute_Edward_distance.R` for code

## Isolation by Distance

### Common computing

#### Barrier distance

```{r}
water <- ne_states(
  country     = c("united states of america", "mexico"),
  returnclass = "sv"
) |>
  crop(y = ext(-124,-113, 28.5, 38.5)) |>
  rasterize(y = rast(ext(-124, -113, 28.5, 38.5), ncols = 1100, nrows = 1000)) |> 
  subst(from = NA, to = 2) |> 
  subst(from = 1, to = NA) |> 
  subst(from = 2, to = 1)

barrier_distance <- function(from, to) {
  background <- water
  background[from] <- 2
  grid_map <- gridDist(background, target = 2)
  extract(grid_map, to) |> as.numeric()
}
```

#### Topographic distance

```{r}
catalina_df_water <- getNOAA.bathy(
    lon1 = -124, lon2 = -115,
    lat1 = 28,   lat2 = 39,
    resolution = .5
  ) |>
  fortify.bathy() |>
  mutate(z = ifelse(z > 0, NA, z)) |> 
  mutate(z = ifelse(z < 0, -z, z)) |> 
  raster::rasterFromXYZ()
```

```{r}
catalina_df <- getNOAA.bathy(
    lon1 = -124, lon2 = -115,
    lat1 = 28,   lat2 = 39,
    resolution = .5
  ) |> 
  fortify.bathy() |>
  mutate(z = ifelse(z > 0, NA, z)) # Remove landmass data

states <- ne_states(
  country     = c("united states of america", "mexico"),
  returnclass = "sf"
)
```

### All populations

#### Barrier distance

```{r}
geo_dist_cells <- here("data", "individuals_coordinates.csv") |> 
  read_csv2() |> 
  select(c(Localisation, Longitude, Latitude, popmap)) |> 
  distinct() |> 
  rename(pop = popmap, Code = Localisation) |> 
  mutate(
      pop  = as.factor(pop),
      Code = as.factor(Code)
  ) |>
  arrange(pop) |> 
  (\(x) {
    mutate(x, cell_id = cellFromXY(water, as.matrix(x[, 2:3])))
  })() |> 
  pull(cell_id, name = pop)
```

```{r}
barrier_dist_mat <- proxy::dist(geo_dist_cells, method = barrier_distance)
```

```{r}
mantel_plot(IBD$Dgeo, barrier_dist_mat)
```

Distances correlate -> don't use barrier distance

#### Geographic distance considering bathymetry

```{r}
coordinates <- here("data", "individuals_coordinates.csv") |>
  read_csv2() |>
  select(popmap, Longitude, Latitude) |> 
  distinct() |> 
  mutate(popmap = as.factor(popmap)) |> 
  arrange(popmap) |> 
  column_to_rownames(var = "popmap") |> 
  as.matrix()
```

```{r}
Topo_Dist_w <- topoDist(catalina_df_water, coordinates, paths = TRUE)
```

```{r}
bathy_paths <- ggplot() +
  geom_raster(data = catalina_df, mapping = aes(x = x, y = y, fill = z)) +
  scale_fill_hypso_tint_c(palette = "colombia_bathy", limits = range(catalina_df$z, na.rm = TRUE)) +
  geom_sf(data = states) +
  gg(data = Topo_Dist_w[[2]]) +
  geom_point(data = coordinates, mapping = aes(x = Longitude, y = Latitude), size = 2) +
  coord_sf(xlim = c(-124, -115), ylim = c(28, 39), expand = FALSE) +
  labs(x = "Longitude", y = "Latitude")

save_open_plot(
  here("output", "plots", "IBD_bathy_map.png"),
  plot   = bathy_paths,
  width  = 9,
  height = 7
)
```

#### Load PCA distances

```{r}
pcadapt_lcwgs <- read_rds(here("output", "pcadapt_lcwgs.rds"))
popmap_lcwgs <- popmap_lcwgs <- c(
  rep("bodega_bay",            5), # BB
  rep("big_creek",            13), # BIGC -> BCR
  rep("catalina_island",      11), # CAT
  rep("elkhorn",               6), # ELK
  rep("guadalupe_island",     10), # GUA
  rep("pacific_grove",        13), # HOP -> Pacific Grove PGR 
  rep("isla_san_jeronimo",    11), # ISJ
  rep("laguna_beach",         13), # LB
  rep("point_dume",            5), # PD
  rep("santa_barbara",        13), # SB
  rep("san_clemente_island",  10), # SCL
  rep("la_jolla_san_diego",   13), # SD
  rep("tomales_bay",           1), # TB
  rep("redondo_beach",        10), # RB
  rep("santa_cruz_harbour",   10) # SCH
)
```

```{r}
PCA_distance <- tibble(
  PCA1 = pcadapt_lcwgs$scores[, 1],
  PCA2 = pcadapt_lcwgs$scores[, 2],
  PCA3 = pcadapt_lcwgs$scores[, 3],
  PCA4 = pcadapt_lcwgs$scores[, 4],
  pop  = popmap_lcwgs
) |> 
  group_by(pop) |> 
  summarise(
    PCA1_popmean = mean(PCA1),
    PCA2_popmean = mean(PCA2),
    PCA3_popmean = mean(PCA3),
    PCA4_popmean = mean(PCA4)
  ) |>
  column_to_rownames(var = "pop") |> 
  dist()

rm(pcadapt_lcwgs, popmap_lcwgs)
```

#### Load Edward distance

```{r}
Edward_distance <- readRDS(here("output", "EJA_lcwgs_Edward_distance.rds"))
```

#### All comparisons

```{r}
mantel.randtest(dist(coordinates), PCA_distance)
p1 <- mantel_plot(as.dist(Topo_Dist_w[[1]]), PCA_distance, xlab = "Euclidean", ylab = "PCA")

mantel.randtest(as.dist(Topo_Dist_w[[1]]), PCA_distance)
p2 <- mantel_plot(as.dist(Topo_Dist_w[[1]]), PCA_distance, xlab = "Bathy", ylab = "PCA")

mantel.randtest(as.dist(geosphere::distm(coordinates)), Edward_distance)
p3 <- mantel_plot(
  as.dist(geosphere::distm(coordinates)),
  Edward_distance,
  xlab = "Ellipsoid Distance", ylab = " Cavalli-Sforza and Edward Chord Distance"
)

mantel.randtest(as.dist(Topo_Dist_w[[1]]), Edward_distance)
p4 <- mantel_plot(as.dist(Topo_Dist_w[[1]]), Edward_distance, xlab = "Bathy", ylab = "Chord distance")

combined <- (p1 | p2) / (p3 | p4)

save_open_plot(
  path   = here("output", "plots", "IBD_all_pops_K3.png"),
  plot   = combined,
  width  = 8,
  height = 8
)
```

```{r}
rm(combined, p1, p2, p3, p4, p5, p6)
rm(Topo_Dist_w, PCA_distance)
```

#### Adegenet analysis

```{r}
Dgen <- Edward_distance
Dgeo <- dist(coordinates)

ibd <- mantel.randtest(Dgen, Dgeo)
vegan::mantel(Dgen, Dgeo)
ibd
plot(ibd)
```

### Coast + ISJ

#### Euclidean distance
```{r}
Dist_geo_isl <- here("data", "individuals_coordinates.csv") |> 
  read_csv2() |> 
  select(c(Localisation, Longitude, Latitude, popmap)) |>
  distinct() |> 
  rename(pop = popmap, Code = Localisation) |> 
  arrange(pop) |> 
  filter(
      !pop %in% c("guadalupe_island", "catalina_island", "san_clemente_island")
  ) |>
  as.data.frame()

rownames(Dist_geo_isl) <- Dist_geo_isl$pop

Dist_geo_isl <- Dist_geo_isl[, 2:3] |> dist()
```

#### Barrier distance

```{r}
coordinates_coast_isj <- here("data", "individuals_coordinates.csv") |> 
  read_csv2() |> 
  select(c(popmap, Longitude, Latitude)) |>
  distinct() |> 
  mutate(Popmap = as.factor(popmap)) |> 
  arrange(Popmap) |> 
  filter(
    !Popmap %in% c("guadalupe_island", "catalina_island", "san_clemente_island")
  ) |> 
  column_to_rownames(var = "Popmap") |> 
  select(-c(popmap)) |> 
  as.matrix()
```

```{r}
vegan::mantel(Dist_geo_isl, lcwgs_IBD_islanders$Dgeo)
```

#### Geographic distance considering bathymetry

```{r}
Topo_Dist_w_coast_isj <- topoDist(catalina_df_water, coordinates_coast_isj, paths = TRUE)
```

```{r}
bathy_paths_coast <- ggplot() +
  geom_raster(data = catalina_df, mapping = aes(x = x, y = y, fill = z)) +
  scale_fill_hypso_tint_c(palette = "colombia_bathy", limits = range(catalina_df$z, na.rm = TRUE)) +
  geom_sf(data = states) +
  gg(data = Topo_Dist_w_coast[[2]]) +
  geom_point(data = coordinates_coast, mapping = aes(x = Longitude, y = Latitude), size = 2) +
  coord_sf(xlim = c(-124, -115), ylim = c(28, 39), expand = FALSE) +
  labs(x = "Longitude", y = "Latitude")

save_open_plot(
  here("output", "plots", "IBD_bathy_map_coast.png"),
  plot   = bathy_paths_coast,
  width  = 9,
  height = 7
)
```

#### Load PCA distance

```{r}
pcadapt_lcwgs_coast_isj <- read_rds(here("output", "pcadapt_lcwgs_coast_isj_K6.rds"))
popmap_lcwgs_coast_isj <- c(
  rep("bodega_bay",            5), # BB
  rep("big_creek",            13), # BIGC -> BCR
  rep("elkhorn",               6), # ELK
  rep("pacific_grove",        13), # HOP -> Pacific Grove PGR
  rep("isla_san_jeronimo",    11), # ISJ
  rep("laguna_beach",         13), # LB
  rep("point_dume",            5), # PD
  rep("santa_barbara",        13), # SB
  rep("la_jolla_san_diego",   13), # SD
  rep("tomales_bay",           1), # TB
  rep("redondo_beach",        10), # RB
  rep("santa_cruz_harbour",   10)  # SCH
)

PCA_distances_coast_isj <- tibble(
  PCA1 = pcadapt_lcwgs_coast_isj$scores[, 1],
  PCA2 = pcadapt_lcwgs_coast_isj$scores[, 2],
  pop  = popmap_lcwgs_coast_isj
) |> 
  group_by(pop) |> 
  summarise(
    PCA1_popmean = mean(PCA1),
    PCA2_popmean = mean(PCA2)
  ) |> 
  column_to_rownames(var = "pop") |> 
  dist()

rm(pcadapt_lcwgs_coast_isj, popmap_lcwgs_coast_isj)
```

#### Load Edward

```{r}
pops_coast_isj <- c("big_creek",         "bodega_bay",         "elkhorn",
                    "isla_san_jeronimo", "la_jolla_san_diego", "laguna_beach",
                    "pacific_grove",     "point_dume",         "redondo_beach",
                    "santa_barbara",     "santa_cruz_harbour", "tomales_bay")

Edward_distance_coast_isj <- here("output", "EJA_lcwgs_Edward_distance.rds") |> 
  readRDS() |>
  as.matrix() |> 
  _[pops_coast_isj, pops_coast_isj] |> 
  as.dist()
```

#### All comparisons

```{r}
mantel.randtest(dist(coordinates_coast_isj), PCA_distances_coast_isj)
p1 <- mantel_plot(dist(coordinates_coast_isj), PCA_distances_coast_isj, xlab = "Euclidean", ylab = "PCA")

mantel.randtest(as.dist(Topo_Dist_w_coast_isj[[1]]), PCA_distances_coast_isj)
p2 <- mantel_plot(as.dist(Topo_Dist_w_coast_isj[[1]]), PCA_distances_coast_isj, xlab = "Bathy", ylab = "PCA")

mtest <- mantel.randtest(as.dist(geosphere::distm(coordinates_coast_isj)), Edward_distance_coast_isj)
p3 <- mantel_plot(
  as.dist(geosphere::distm(coordinates_coast_isj, fun = 
geosphere::distVincentyEllipsoid)),
  Edward_distance_coast_isj,
  xlab = "Ellipsoid Distance",
  ylab = " Cavalli-Sforza and Edward Chord Distance"
)

mantel.randtest(as.dist(Topo_Dist_w_coast_isj[[1]]), Edward_distance_coast_isj)
p4 <- mantel_plot(as.dist(Topo_Dist_w_coast_isj[[1]]), Edward_distance_coast_isj, xlab = "Bathy", ylab = "Chord distance")

combined <- (p1 | p2) / (p3 | p4)

save_open_plot(
  path   = here("output", "plots", "IBD_coast_isj_pops.png"),
  plot   = combined,
  width  = 8,
  height = 8
  )

save_open_plot(
  path   = here("output", "plots", "IBD_coast_isj_Euclidean_Edward.pdf"),
  plot   = p3,
  width  = 6,
  height = 6
  )
```

```{r}
rm(combined, p1, p2, p3, p4)
```

### Coast

#### Euclidean distance
```{r}
coordinates_coast <- here("data", "individuals_coordinates.csv") |>
  read_csv2() |>
  select(popmap, Longitude, Latitude) |>
  distinct() |> 
  mutate(Popmap = as.factor(popmap)) |> 
  arrange(Popmap) |> 
  filter(
    !Popmap %in% c("guadalupe_island", "catalina_island", "san_clemente_island", "isla_san_jeronimo")
  ) |> 
  column_to_rownames(var = "Popmap") |> 
  as.matrix()
```

#### Geographic distance considering bathymetry

```{r}
Topo_Dist_w_coast <- topoDist(catalina_df_water, coordinates_coast, paths = TRUE)
```

```{r}
bathy_paths_coast <- ggplot() +
  geom_raster(data = catalina_df, mapping = aes(x = x, y = y, fill = z)) +
  scale_fill_hypso_tint_c(palette = "colombia_bathy", limits = range(catalina_df$z, na.rm = TRUE)) +
  geom_sf(data = states) +
  gg(data = Topo_Dist_w_coast[[2]]) +
  geom_point(data = coordinates_coast, mapping = aes(x = Longitude, y = Latitude), size = 2) +
  coord_sf(xlim = c(-124, -115), ylim = c(28, 39), expand = FALSE) +
  labs(x = "Longitude", y = "Latitude")

save_open_plot(
  here("output", "plots", "IBD_bathy_map_coast.png"),
  plot   = bathy_paths_coast,
  width  = 9,
  height = 7
)
```

#### LoadPCA distance

```{r}
pcadapt_lcwgs_coast <- read_rds(here("output", "pcadapt_lcwgs_coast_K5.rds"))
popmap_lcwgs_coast <- c(
  rep("bodega_bay",            5), # BB
  rep("big_creek",            13), # BIGC -> BCR
  rep("elkhorn",               6), # ELK
  rep("pacific_grove",        13), # HOP -> Pacific Grove PGR
  rep("laguna_beach",         13), # LB
  rep("point_dume",            5), # PD
  rep("santa_barbara",        13), # SB
  rep("la_jolla_san_diego",   13), # SD
  rep("tomales_bay",           1), # TB
  rep("redondo_beach",        10), # RB
  rep("santa_cruz_harbour",   10)  # SCH
)

PCA_distances_coast <- tibble(
  PCA1 = pcadapt_lcwgs_coast$scores[, 1],
  PCA2 = pcadapt_lcwgs_coast$scores[, 2],
  pop  = popmap_lcwgs_coast
) |> 
  group_by(pop) |> 
  summarise(
    PCA1_popmean = mean(PCA1),
    PCA2_popmean = mean(PCA2)
  ) |> 
  column_to_rownames(var = "pop") |> 
  dist()

rm(pcadapt_lcwgs_coast, popmap_lcwgs_coast)
```

#### Load Edward distance

```{r}
pops_coast <-  c("big_creek",         "bodega_bay",         "elkhorn",
                 "la_jolla_san_diego", "laguna_beach", "pacific_grove",
                 "point_dume",         "redondo_beach", "santa_barbara",
                 "santa_cruz_harbour", "tomales_bay")

Edward_distance_coast <- here("output", "EJA_lcwgs_Edward_distance.rds") |> 
  readRDS() |>
  as.matrix() |> 
  _[pops_coast, pops_coast] |> 
  as.dist()
```

#### All comparisons

```{r}
mantel.randtest(dist(coordinates_coast), PCA_distances_coast) |> plot()
p1 <- mantel_plot(dist(coordinates_coast),
                  PCA_distances_coast,
                  xlab = "Euclidean", ylab = "PCA")

mantel.randtest(as.dist(Topo_Dist_w_coast[[1]]), PCA_distances_coast) |> plot()
p2 <- mantel_plot(as.dist(Topo_Dist_w_coast[[1]]),
                  PCA_distances_coast,
                  xlab = "Bathy", ylab = "PCA")

mantel.randtest(dist(coordinates_coast), Edward_distance_coast) |> plot()
p3 <- mantel_plot(as.dist(geosphere::distm(coordinates_coast)),
                  Edward_distance_coast,
                  xlab = "Euclidean", ylab = "Edward Chord distance")

mantel.randtest(as.dist(Topo_Dist_w_coast[[1]]), Edward_distance_coast) |> plot()
p4 <- mantel_plot(as.dist(Topo_Dist_w_coast[[1]]),
                  Edward_distance_coast,
                  xlab = "Bathy", ylab = "Edward Chord distance")

combined <- (p1 | p2) / (p3 | p4)

save_open_plot(
  path   = here("output", "plots", "IBD_coast_Euclidean_Edward.png"),
  plot   = p3,
  width  = 8,
  height = 8
)

save_open_plot(
  path   = here("output", "plots", "IBD_coast_pops.png"),
  plot   = combined,
  width  = 8,
  height = 8
  )
```

```{r}
rm(combined, p1, p2, p3, p4, p5, p6)
```

### Tree tests

```{r}
IBD$Dgen |> nj() |> root(outgroup = "guadalupe_island") |> ladderize() |> plot()
```

```{r}
PCA_distance <- tibble(
  PCA1 = pcadapt_lcwgs$scores[, 1],
  PCA2 = pcadapt_lcwgs$scores[, 2],
  PCA3 = pcadapt_lcwgs$scores[, 3],
  pop  = popmap_lcwgs
) |> 
  group_by(pop) |> 
  summarise(
    PCA1_popmean = mean(PCA1),
    PCA2_popmean = mean(PCA2),
    PCA3_popmean = mean(PCA3)
  ) |>
  column_to_rownames(var = "pop") |> 
  dist()
```

```{r}
PCA_distance |> nj() |> root(outgroup = "guadalupe_island") |> ladderize() |> plot()
```

```{r}
Edward_distance |> nj() |> root(outgroup = "guadalupe_island") |> ladderize() |> plot()
```

## Admixture analysis

### Convert data to ped format for LEA import using plink2
```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2
./plink2 --vcf 8-Embiotoca_filtered.vcf.gz --export ped --out 8-Embiotoca_filtered
```

### Read and convert data
```{r}
#| label: VcfConvert
#| eval: false

ped2geno(
  input.file  = here("plink2", "8-Embiotoca_filtered.ped"),
  output.file = here("data", "SNPs", "8-Embiotoca_filtered.geno")
)
```

### Run analysis
```{r}
#| label: snmfAnalysisSingle
#| eval: false

snmf_analysis <- snmf(
  here("data", "SNPs", "8-Embiotoca_filtered.geno"),
  K           = 4:9,
  project     = "continue",
  repetitions = 50,
  entropy     = TRUE
)
```

### Load analysis

```{r}
EJA_snmf_analysis <- load.snmfProject(here("data", "lcwgs", "8-Embiotoca_filtered.snmfProject"))
```

#### Plot all cross-entropies to check distribution

```{r}
cross_entropies <- tibble(
  `c-e`   = map(4:9, \(x) cross.entropy(EJA_snmf_analysis, K = x) |>
                            as.vector()) |> unlist(),
  K_value = factor(c(rep(4, 50), rep(5, 50), rep(6, 50),
                     rep(7, 50), rep(8, 50), rep(9, 10)))
)

cross_entropies |> 
  ggplot(aes(x = K_value, y = `c-e`)) +
  geom_boxplot(alpha = .3) +
  # geom_violin(alpha = .3) +
  geom_point(alpha = .1) +
  labs(x = "Number of ancestral population", y = "Cross-entropy Criterion") +
  theme_classic()

cross_entropies |>
  filter(K_value == 7) |>
  mutate(run = row_number()) |> 
  ggplot() +
  geom_point(aes(x = run, y = `c-e`))

rm(cross_entropies)
```

### Plot minimal cross entropy to chose optimal ancestral population amount

```{r}
#| label: snmfCrossEntropy

plot(snmf_analysis)

snmf_tb <- tibble(K = as.numeric(unique(snmf_analysis@K)))

cross_entropy <- rep(0, dim(snmf_tb)[1])

for (i in 1:length(snmf_tb$K)) {
  cross_entropy[i] <- min(cross.entropy(snmf_analysis, K = snmf_tb$K[i]))
}

snmf_plot <- snmf_tb |>
  mutate(cross_entropy = cross_entropy) |> 
  ggplot() +
  geom_point(aes(x = K, y = cross_entropy)) +
  theme_classic()

ggsave(
  filename = here("output", "plots", "snmf_cross_entropy.png"),
  plot     = snmf_plot,
  width    = 8,
  height   = 5
)
```

K = 7 has the minimum cross-entropy.

#### Select best run to plot for each K value

```{r}
which.min(cross.entropy(EJA_snmf_analysis, K = 5))
which.min(cross.entropy(EJA_snmf_analysis, K = 6))
which.min(cross.entropy(EJA_snmf_analysis, K = 7))
which.min(cross.entropy(EJA_snmf_analysis, K = 8))
which.min(cross.entropy(EJA_snmf_analysis, K = 9))
```

### Plot results for K = 7

```{r}
#| label: snmfBarPlot

barchart(
  snmf_analysis,
  K         = 7,
  run       = which.min(cross.entropy(snmf_analysis, K = 7)),
  col       = viridis::turbo(n = 6),
  sort.by.Q = FALSE
)
```

See `EJA_populations_lcwgs_Figures.qmd` for pretty plots.

## Triangle plots

### Coastal

```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

./plink2 --bfile 8-Embiotoca_filtered_coastal --keep-allele-order --recode vcf --out 8-Embiotoca_filtered_coastal
```

```{r}
lcwgs_coast_vcf <- read.vcfR(here("plink2", "8-Embiotoca_filtered_coastal.vcf"))

popmap_lcwgs_coast <- read_csv2(here("data", "individuals_coordinates.csv")) |> 
  select(Indiv, Localisation) |>
  rename(id = Indiv, pop = Localisation) |> 
  filter(pop != c("CAT")) |> 
  filter(pop != c("ISJ")) |>
  filter(pop != c("SCL")) |> 
  filter(pop != c("GUA")) |> 
  data.frame()
```

```{r}
lcwgs_coast_vcf_diffs <- vector(mode = 'list', length = 3)
parameters <- c(.5, .75, 1)

for (i in 1:3) {
  lcwgs_coast_vcf_diffs[[i]] <- alleleFreqDiff(lcwgs_coast_vcf,
                                               pm = popmap_lcwgs_coast,
                                               p1 = "BBA", p2 = "SD",
                                               difference = parameters[i])
}

rm(lcwgs_coast_vcf)
```

with 1 difference : "8456 sites passed allele frequency difference threshold"

Difference = .75 76178

Difference = .5 : 276405 sites

```{r}
hi_het_coast <- vector(mode = 'list', length = 3)

for (i in 1:3) {
  hi_het_coast[[i]] <- hybridIndex(lcwgs_coast_vcf_diffs[[i]],
                                   pm = popmap_lcwgs_coast,
                                   p1 = "BBA", p2 = "SD")
}
```

```{r}
#| eval: false

write_rds(hi_het_coast, here("output", "Triangle_analysis_coast.rds"))
```

```{r}
hi_het_coast_plots <- vector(mode = 'list', length = 3)

for (i in 1:3) {
  hi_het_coast_plots[[i]] <- hi_het_coast[[i]] |>
    mutate(
      pop = factor(
        pop,
        levels = c("BBA", "TB", "SCH", "ELK", "HOP",
                   "BIGC", "SB", "PD", "RB", "LAG", "SD")
      )
    ) |> 
    arrange(pop) |>
    triangle.plot(colors = turbo(n = 11)) +
    labs(title = paste0("d = ", as.character(parameters[i])))
}

save_open_plot(
  path   = here("output", "plots", "coast_Triangle_plots.pdf"),
  plot   = Reduce("+", hi_het_coast_plots) +
    plot_layout(axis_titles = "collect", guides = "collect"),
  width  = 14,
  height = 4
)
```

```{r}
rm(lcwgs_coast_vcf_diffs, hi_het_coast, hi_het_coast_plots)
gc()
```

### Coastal + ISJ

```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

./plink2 --bfile 8-Embiotoca_filtered_coastal_isj --keep-allele-order --recode vcf --out 8-Embiotoca_filtered_coastal_isj
```

```{r}
EJA_lcwgs_coast_isj_vcf <- read.vcfR(here("plink2", "8-Embiotoca_filtered_coastal_isj.vcf"))

popmap_lcwgs_coast_isj <- read_csv2(here("data", "individuals_coordinates.csv")) |> 
  select(Indiv, Localisation) |>
  rename(id = Indiv, pop = Localisation) |> 
  filter(pop != c("CAT")) |> 
  filter(pop != c("SCL")) |> 
  filter(pop != c("GUA")) |> 
  data.frame()
```

```{r}
EJA_lcwgs_coast_isj_vcf_diff <- vector(mode = 'list', length = 3)
parameters <- c(.5, .75, 1)

for (i in 1:3) {
  EJA_lcwgs_coast_isj_vcf_diff[[i]] <- alleleFreqDiff(EJA_lcwgs_coast_isj_vcf,
                                                      pm = popmap_lcwgs_coast_isj,
                                                      p1 = "BBA", p2 = "ISJ",
                                                      difference = parameters[i])
}

rm(EJA_lcwgs_coast_isj_vcf)
gc()
```

With 1 : "32484 sites passed allele frequency difference threshold"
With .75 : "166077 sites passed allele frequency difference threshold"
With .5 : "424389 sites passed allele frequency difference threshold"

```{r}
hi_het_coast_isj <- vector(mode = 'list', length = 3)

for (i in 1:3) {
  hi_het_coast_isj[[i]] <- hybridIndex(EJA_lcwgs_coast_isj_vcf_diff[[i]],
                                       pm = popmap_lcwgs_coast_isj,
                                       p1 = "BBA", p2 = "ISJ")
}
```

```{r}
#| eval: false

write_rds(hi_het_coast_isj, here("output", "Triangle_analysis_coast_isj.rds"))
```

```{r}
hi_het_coast_isj_plots <- vector(mode = 'list', length=3)

for (i in 1:3) {
  hi_het_coast_isj_plots[[i]] <- hi_het_coast_isj[[i]] |>
    mutate(
      pop = factor(
        pop,
        levels = c("BBA", "TB", "SCH", "ELK", "HOP", "BIGC",
                   "SB",  "PD", "RB",  "LAG", "SD",  "ISJ")
      )
    ) |> 
    arrange(pop) |>
    triangulaR::triangle.plot(colors = turbo(n = 12)) +
    labs(title = paste0("d = ", as.character(parameters[i])))
}

save_open_plot(
  path   = here("output", "plots", "coast_isj_Triangle_plots.pdf"),
  plot   = Reduce("+", hi_het_coast_isj_plots) +
    plot_layout(axis_titles = "collect", guides = "collect"),
  width  = 14,
  height = 4
)
```

```{r}
rm(EJA_lcwgs_coast_isj_vcf_diff, hi_het_coast_isj, hi_het_coast_isj_plots)
gc()
```

### North

```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

./plink2 --bfile 8-Embiotoca_filtered_north --keep-allele-order --recode vcf --out 8-Embiotoca_filtered_north
```

```{r}
EJA_lcwgs_north_vcf <- read.vcfR(here("plink2", "8-Embiotoca_filtered_north.vcf"))

popmap_lcwgs_north <- read_csv2(here("data", "individuals_coordinates.csv")) |> 
  select(Indiv, Localisation) |>
  rename(id = Indiv, pop = Localisation) |> 
  filter(pop %in% c("BBA", "BIGC", "ELK", "HOP", "SB", "TB", "SCH")) |> 
  data.frame()
```

```{r}
EJA_lcwgs_north_vcf_diff <- vector(mode='list', length=3)
parameters <- c(.5, .75, 1)

for (i in 1:3) {
  EJA_lcwgs_north_vcf_diff[[i]] <- alleleFreqDiff(EJA_lcwgs_north_vcf,
                                                  pm = popmap_lcwgs_north,
                                                  p1 = "BBA", p2 = "SB",
                                                  difference = parameters[i])
}

rm(EJA_lcwgs_north_vcf)
gc()
```

Difference = 1.0 : 2348 sites
Difference = 0.9 : 9364 sites
Difference = 0.8 : 26991 sites (>1% SNPs)
Difference = 0.75 : 38762
Difference = 0.7 : 55736 sites
Difference = 0.5 : 183805 sites

```{r}
hi_het_north <- vector(mode='list', length=3)

for (i in 1:3) {
  hi_het_north[[i]] <- hybridIndex(EJA_lcwgs_north_vcf_diff[[i]],
                                   pm = popmap_lcwgs_north,
                                   p1 = "BBA", p2 = "SB")
}
```

```{r}
#| eval: false

write_rds(hi_het_north, here("output", "Triangle_analysis_north.rds"))
```

```{r}
hi_het_north_plots <- vector(mode='list', length=3)

for (i in 1:3) {
  hi_het_north_plots[[i]] <- hi_het_north[[i]] |>
    mutate(
      pop = factor(
        pop,
        levels = c("BBA", "TB", "SCH", "ELK", "HOP", "BIGC", "SB")
      )
    ) |> 
    arrange(pop) |>
    triangle.plot(colors = turbo(n = 7)) +
    labs(title = paste0("d = ", as.character(parameters[i])))
}

save_open_plot(
  path   = here("output", "plots", "north_Triangle_plots.pdf"),
  plot   = Reduce("+", hi_het_north_plots) +
    plot_layout(axis_titles = "collect", guides = "collect"),
  width  = 14,
  height = 4
)
```


```{r}
rm(EJA_lcwgs_north_vcf_diff, hi_het_north, hi_het_north_plots)
gc()
```

### South

```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

./plink2 --bfile 8-Embiotoca_filtered_south --keep-allele-order --recode vcf --out 8-Embiotoca_filtered_south
```

```{r}
EJA_lcwgs_south_vcf <- read.vcfR(here("plink2", "8-Embiotoca_filtered_south.vcf"))

popmap_lcwgs_south <- read_csv2(here("data", "individuals_coordinates.csv")) |> 
  select(Indiv, Localisation) |>
  rename(id = Indiv, pop = Localisation) |> 
  filter(pop %in% c("SB", "PD", "RB", "LAG", "SD")) |> 
  data.frame()
```

```{r}
EJA_lcwgs_south_vcf_diff <- vector(mode='list', length=3)
parameters <- c(.5, .75, 1)

for (i in 1:3) {
  EJA_lcwgs_south_vcf_diff[[i]] <- alleleFreqDiff(EJA_lcwgs_south_vcf,
                                                  pm = popmap_lcwgs_south,
                                                  p1 = "SB", p2 = "SD",
                                                  difference = parameters[i])
}

rm(EJA_lcwgs_south_vcf)
gc()
```

With 1 : 234
.75 : 3367
.5 : 54787

```{r}
hi_het_south <- vector(mode='list', length=3)

for (i in 1:3) {
  hi_het_south[[i]] <- hybridIndex(EJA_lcwgs_south_vcf_diff[[i]],
                                   pm = popmap_lcwgs_south,
                                   p1 = "SB", p2 = "SD")
}
```

```{r}
#| eval: false

write_rds(hi_het_south, here("output", "Triangle_analysis_south.rds"))
```

```{r}
hi_het_south_plots <- vector(mode='list', length=3)

for (i in 1:3) {
  hi_het_south_plots[[i]] <- hi_het_south[[i]] |>
    mutate(
      pop = factor(
        pop,
        levels = c("SB", "PD", "RB", "LAG", "SD")
      )
    ) |> 
    arrange(pop) |>
    triangle.plot(colors = turbo(n = 5)) +
    labs(title = paste0("d = ", as.character(parameters[i])))
}

save_open_plot(
  path   = here("output", "plots", "south_Triangle_plots.pdf"),
  plot   = Reduce("+", hi_het_south_plots) +
    plot_layout(axis_titles = "collect", guides = "collect"),
  width  = 14,
  height = 4
)
```

```{r}
rm(EJA_lcwgs_south_vcf_diff, hi_het_south, hi_het_south_plots)
gc()
```

## Genotype-Environment Association

<!-- ## Prep -->

<!-- ```{r} -->
<!-- #| eval: false -->

<!-- Temperatures <- read_rds(here("data", "Copernicus", "annual_average_SST.rds")) |>  -->
<!--   terra::unwrap() -->

<!-- Coordinates <- read_csv2(here("data", "individuals_coordinates_temps.csv")) |> -->
<!--   select(Longitude, Latitude) |>  -->
<!--   as.data.frame() -->

<!-- Temp_data <- terra::extract( -->
<!--   x = Temperatures, -->
<!--   y = data.frame(Coordinates), -->
<!--   cells = FALSE -->
<!-- ) -->

<!-- read_csv2(here("data", "individuals_coordinates_temps.csv")) |>  -->
<!--   mutate("Temp (Celsius)" = Temp_data$mean) |>  -->
<!--   write_csv2(here("data", "individuals_coordinates_temps.csv")) -->

<!-- rm(Temperatures, Coordinates, Temp_data) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- ggplot() + -->
<!--   geom_spatraster(data = Temperatures) + -->
<!--   geom_point(data = Coordinates, aes(x = Longitude, y = Latitude), shape = 21) -->
<!-- ``` -->

<!-- ## All pops -->

<!-- ### Data preparation -->

<!-- ```{r} -->
<!-- Temp_data <- read_csv2(here("data", "individuals_coordinates_temps.csv")) |>  -->
<!--   pull(`Temp (Celsius)`) |>  -->
<!--   as.matrix() -->
<!-- ``` -->


<!-- ```{r} -->
<!-- Genotype <- LEA::read.geno(here("data", "8-Embiotoca_filtered.geno")) -->

<!-- Genotype_noNA <- Genotype[, !apply(Genotype, 2, function(col) any(col == 9))] -->

<!-- snmf(here("data", "8-Embiotoca_filtered.geno"), K = 7,  project = "new",) -->

<!-- impute( -->
<!--   object     = load.snmfProject( -->
<!--     here("data", "snmf", "8-Embiotoca_filtered.snmfProject") -->
<!--   ), -->
<!--   input.file = here("data", "8-Embiotoca_filtered.geno"), -->
<!--   method      = "mode", -->
<!--   K          = 7, -->
<!--   run        = 1 -->
<!-- ) -->

<!-- Genotype_imputed <- read.lfmm(here("data", "8-Embiotoca_filtered.lfmm_imputed.lfmm")) -->

<!-- rm(Genotype) -->
<!-- ``` -->

<!-- ### Analysis -->

<!-- ```{r} -->
<!-- mod_lfmm <- LEA::lfmm2( -->
<!--   input        = Genotype_noNA, -->
<!--   env          = Temp_data,  -->
<!--   K            = 4,  -->
<!--   effect.sizes = TRUE -->
<!-- ) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- pv <-  lfmm2.test(mod_lfmm,  -->
<!--                   input = Genotype_noNA,  -->
<!--                   env   = Temp_data,  -->
<!--                   full  = TRUE) -->

<!-- plot(-log10(pv$pvalue), cex = .3, pch = 19, col = "blue") -->
<!-- ``` -->

<!-- ## Coast -->

<!-- ### Data preparation -->

<!-- ```{r} -->
<!-- #| eval: false -->

<!-- vcf2geno( -->
<!--   input.file  = here("plink2", "8-Embiotoca_filtered_coastal.vcf"), -->
<!--   output.file = here("data", "8-Embiotoca_filtered_coastal.geno") -->
<!-- ) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- Genotype_coast <- LEA::read.geno(here("data", "8-Embiotoca_filtered_coastal.geno")) -->

<!-- Genotype_coast_noNA <- Genotype_coast[, !apply(Genotype_coast, 2, function(col) any(col == 9))] -->

<!-- rm(Genotype_coast) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- Temp_data_coast <- read_csv2(here("data", "individuals_coordinates_temps.csv")) |>  -->
<!--   filter(Localisation != "CAT") |> -->
<!--   filter(Localisation != "GUA") |> -->
<!--   filter(Localisation != "ISJ") |> -->
<!--   filter(Localisation != "SCL") |> -->
<!--   pull(`Temp (Celsius)`) |>  -->
<!--   as.matrix() -->
<!-- ``` -->


<!-- ### Analysis -->

<!-- ```{r} -->
<!-- mod_lfmm_coast <- LEA::lfmm2( -->
<!--   input        = Genotype_coast_noNA, -->
<!--   env          = Temp_data_coast,  -->
<!--   K            = 3, -->
<!--   effect.sizes = TRUE -->
<!-- ) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- pv_coast <- lfmm2.test( -->
<!--     object = mod_lfmm_coast, -->
<!--     input  = Genotype_coast_noNA,  -->
<!--     env    = Temp_data_coast,  -->
<!--     full   = TRUE -->
<!--   ) -->

<!-- plot(-log10(pv_coast$pvalue), cex = .3, pch = 19, col = "blue") -->
<!-- ``` -->
