---
title: "EJA Population Genomics Low coverage"
author: "Arthur Oulès"
format: html
editor: source
engine: knitr
---
# Load packages

```{r}
#| label: LoadPackages
#| echo: false

c("here", "tidyverse",
  "pcadapt", "vcfR", "adegenet", "ape", "LEA", "triangulaR",
  "viridisLite") |>
  lapply(library, character.only = TRUE) |> invisible()
```

# Load data analysis functions
```{r}
#| label: LoadExtFunc

source(here("functions", "lcwgs_tidy_functions.R"))
source(here("functions", "plot_functions.R"))
```

# Helper objects

```{r}
#| label: GenomeMap
#| echo: false

lcwgs_vcf <- read.vcfR(here("data", "SNPS", "8-Embiotoca_filtered.vcf.gz"))

lcwgs_vcf |>
  getFIX() |>
  as_tibble() |>
  select(CHROM, POS) |>
  mutate(POS = as.numeric(POS)) |>
  mutate(rank = row_number()) |>
  select(rank, CHROM, POS) |> 
  saveRDS(file = here("data", "SNP_positions.rds"))

lcwgs_vcf |>
  extract_gt_tidy() |>
  _$Indiv |>
  write_rds(file = here("data", "Individual_list.rds"))

rm(lcwgs_vcf)
```

# pcadapt analyses

## Convert vcf to bed file using plink2

### All pops
```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

./plink2 --vcf 8-Embiotoca_filtered.vcf.gz --make-bed --out 8-Embiotoca_filtered
```

### Generate whitelists and subset bed files

```{r}
#| label: CoastSaveWhitelist
#| eval: false

read_csv2(here("data", "individuals_coordinates.csv")) |>
  pull(Indiv) |> 
  {\(x) {
    x[!(grepl(pattern = "SCL", x = x) |
        grepl(pattern = "GUA", x = x) |
        grepl(pattern = "CAT", x = x) |
        grepl(pattern = "ISJ", x = x))]
    }}() |> 
  write_lines(here("plink2", "whitelist_coastal.txt"))
```

```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

./plink2 --bfile 8-Embiotoca_filtered --keep whitelist_coastal.txt --make-bed --out 8-Embiotoca_filtered_coastal
```

```{r}
#| label: ISLSaveWhitelist
#| eval: false

read_csv2(here("data", "individuals_coordinates.csv")) |>
  pull(Indiv) |> 
  {\(x) {
    x[(grepl(pattern = "CAT", x = x) |
       grepl(pattern = "GUA", x = x) |
       grepl(pattern = "ISJ", x = x) |
       grepl(pattern = "SCL", x = x))]
    }}() |> 
  write_lines(here("plink2", "whitelist_islands.txt"))
```

```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

./plink2 --bfile 8-Embiotoca_filtered --keep whitelist_islands.txt --make-bed --out 8-Embiotoca_filtered_isl
```

```{r}
#| label: CoastISJSaveWhitelist
#| eval: false

read_csv2(here("data", "individuals_coordinates.csv")) |>
  pull(Indiv) |> 
  {\(x) {
    x[!(grepl(pattern = "SCL", x = x) |
        grepl(pattern = "GUA", x = x) |
        grepl(pattern = "CAT", x = x))]
    }}() |> 
  write_lines(here("plink2", "whitelist_coastal_isj.txt"))
```

```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

./plink2 --bfile 8-Embiotoca_filtered --keep whitelist_coastal_isj.txt --make-bed --out 8-Embiotoca_filtered_coastal_isj
```

```{r}
#| label: CoastSCLSaveWhitelist
#| eval: false

read_csv2(here("data", "individuals_coordinates.csv")) |>
  pull(Indiv) |> 
  {\(x) {
    x[!(grepl(pattern = "ISJ", x = x) |
        grepl(pattern = "GUA", x = x) |
        grepl(pattern = "CAT", x = x))]
    }}() |> 
  write_lines(here("plink2", "whitelist_coastal_scl.txt"))
```

```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

./plink2 --bfile 8-Embiotoca_filtered --keep whitelist_coastal_scl.txt --make-bed --out 8-Embiotoca_filtered_coastal_scl
```

```{r}
#| label: CoastGUASaveWhitelist
#| eval: false

read_csv2(here("data", "individuals_coordinates.csv")) |>
  pull(Indiv) |> 
  {\(x) {
    x[!(grepl(pattern = "ISJ", x = x) |
        grepl(pattern = "SCL", x = x) |
        grepl(pattern = "CAT", x = x))]
    }}() |> 
  write_lines(here("plink2", "whitelist_coastal_gua.txt"))
```

```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

./plink2 --bfile 8-Embiotoca_filtered --keep whitelist_coastal_gua.txt --make-bed --out 8-Embiotoca_filtered_coastal_gua
```

```{r}
#| label: CoastCATSaveWhitelist
#| eval: false

read_csv2(here("data", "individuals_coordinates.csv")) |>
  pull(Indiv) |> 
  {\(x) {
    x[!(grepl(pattern = "ISJ", x = x) |
        grepl(pattern = "SCL", x = x) |
        grepl(pattern = "GUA", x = x))]
    }}() |> 
  write_lines(here("plink2", "whitelist_coastal_cat.txt"))
```

```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

./plink2 --bfile 8-Embiotoca_filtered --keep whitelist_coastal_cat.txt --make-bed --out 8-Embiotoca_filtered_coastal_cat
```

```{r}
#| label: CATSCLSaveWhitelist
#| eval: false

read_csv2(here("data", "individuals_coordinates.csv")) |>
  pull(Indiv) |> 
  {\(x) {
    x[(grepl(pattern = "SCL", x = x) |
       grepl(pattern = "CAT", x = x))]
    }}() |> 
  write_lines(here("plink2", "whitelist_cat_scl.txt"))
```

```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

./plink2 --bfile 8-Embiotoca_filtered --keep whitelist_cat_scl.txt --make-bed --out 8-Embiotoca_filtered_cat_scl
```

```{r}
#| label: GUACATSaveWhitelist
#| eval: false

read_csv2(here("data", "individuals_coordinates.csv")) |>
  pull(Indiv) |> 
  {\(x) {
    x[(grepl(pattern = "GUA", x = x) |
       grepl(pattern = "CAT", x = x))]
    }}() |> 
  write_lines(here("plink2", "whitelist_gua_cat.txt"))
```

```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

./plink2 --bfile 8-Embiotoca_filtered --keep whitelist_gua_cat.txt --make-bed --out 8-Embiotoca_filtered_gua_cat
```

```{r}
#| label: GUASCLSaveWhitelist
#| eval: false

read_csv2(here("data", "individuals_coordinates.csv")) |>
  pull(Indiv) |> 
  {\(x) {
    x[(grepl(pattern = "SCL", x = x) |
       grepl(pattern = "GUA", x = x))]
    }}() |> 
  write_lines(here("plink2", "whitelist_gua_scl.txt"))
```

```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

./plink2 --bfile 8-Embiotoca_filtered --keep whitelist_gua_scl.txt --make-bed --out 8-Embiotoca_filtered_gua_scl
```

```{r}
#| label: GUAISJSaveWhitelist
#| eval: false

read_csv2(here("data", "individuals_coordinates.csv")) |>
  pull(Indiv) |> 
  {\(x) {
    x[(grepl(pattern = "GUA", x = x) |
       grepl(pattern = "ISJ", x = x))]
    }}() |> 
  write_lines(here("plink2", "whitelist_gua_isj.txt"))
```

```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

./plink2 --bfile 8-Embiotoca_filtered --keep whitelist_GUA_ISJ.txt --make-bed --out 8-Embiotoca_filtered_GUA_ISJ
```

```{r}
#| label: CATISJSaveWhitelist
#| eval: false

read_csv2(here("data", "individuals_coordinates.csv")) |>
  pull(Indiv) |> 
  {\(x) {
    x[(grepl(pattern = "CAT", x = x) |
       grepl(pattern = "ISJ", x = x))]
    }}() |> 
  write_lines(here("plink2", "whitelist_cat_isj.txt"))
```

```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

./plink2 --bfile 8-Embiotoca_filtered --keep whitelist_cat_isj.txt --make-bed --out 8-Embiotoca_filtered_cat_isj
```

```{r}
#| label: SCLISJSaveWhitelist
#| eval: false

read_csv2(here("data", "individuals_coordinates.csv")) |>
  pull(Indiv) |> 
  {\(x) {
    x[(grepl(pattern = "ISJ", x = x) |
       grepl(pattern = "SCL", x = x))]
    }}() |> 
  write_lines(here("plink2", "whitelist_scl_isj.txt"))
```

```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

./plink2 --bfile 8-Embiotoca_filtered --keep whitelist_scl_isj.txt --make-bed --out 8-Embiotoca_filtered_scl_isj
```

## All preanalysis to determine optimal K value

```{r}
#| label: PreAnalyses

bed_files <- tibble(paths = list.files(here("plink2")) |> str_subset(".bed")) |> 
  mutate(
    suffixes = paths |>
      str_remove("8-Embiotoca_filtered") |>
      str_remove(".bed") |>
      str_remove("_")
  )

bed_files$suffixes[[1]] <- "all"

bed_files$popmaps <- list(
  c(
    rep("bodega_bay",            5), # BB
    rep("big_creek",            13), # BIGC -> BCR
    rep("catalina_island",      11), # CAT
    rep("elkhorn",               6), # ELK
    rep("guadalupe_island",     10), # GUA
    rep("pacific_grove",        13), # HOP -> Pacific Grove PGR 
    rep("isla_san_jeronimo",    11), # ISJ
    rep("laguna_beach",         13), # LB
    rep("point_dume",            5), # PD
    rep("santa_barbara",        13), # SB
    rep("san_clemente_island",  10), # SCL
    rep("la_jolla_san_diego",   13), # SD
    rep("tomales_bay",           1), # TB
    rep("redondo_beach",        10), # RB
    rep("santa_cruz_harbour",   10) # SCH
  ), c(
    rep("catalina_island",      11), # CAT
    rep("isla_san_jeronimo",    11)  # ISJ
  ), c(
    rep("catalina_island",      11), # CAT
    rep("san_clemente_island",  10)  # SCL
  ), c(
    rep("bodega_bay",            5), # BB
    rep("big_creek",            13), # BIGC -> BCR
    rep("elkhorn",               6), # ELK
    rep("pacific_grove",        13), # HOP -> Pacific Grove PGR
    rep("laguna_beach",         13), # LB
    rep("point_dume",            5), # PD
    rep("santa_barbara",        13), # SB
    rep("la_jolla_san_diego",   13), # SD
    rep("tomales_bay",           1), # TB
    rep("redondo_beach",        10), # RB
    rep("santa_cruz_harbour",   10)  # SCH
  ), c(
    rep("bodega_bay",            5), # BB
    rep("big_creek",            13), # BIGC -> BCR
    rep("catalina_island",      11), # CAT
    rep("elkhorn",               6), # ELK
    rep("pacific_grove",        13), # HOP -> Pacific Grove PGR
    rep("laguna_beach",         13), # LB
    rep("point_dume",            5), # PD
    rep("santa_barbara",        13), # SB
    rep("la_jolla_san_diego",   13), # SD
    rep("tomales_bay",           1), # TB
    rep("redondo_beach",        10), # RB
    rep("santa_cruz_harbour",   10)  # SCH
  ), c(
    rep("bodega_bay",            5), # BB
    rep("big_creek",            13), # BIGC -> BCR
    rep("elkhorn",               6), # ELK
    rep("guadalupe_island",     10), # GUA
    rep("pacific_grove",        13), # HOP -> Pacific Grove PGR
    rep("laguna_beach",         13), # LB
    rep("point_dume",            5), # PD
    rep("santa_barbara",        13), # SB
    rep("la_jolla_san_diego",   13), # SD
    rep("tomales_bay",           1), # TB
    rep("redondo_beach",        10), # RB
    rep("santa_cruz_harbour",   10)  # SCH
  ), c(
    rep("bodega_bay",            5), # BB
    rep("big_creek",            13), # BIGC -> BCR
    rep("elkhorn",               6), # ELK
    rep("pacific_grove",        13), # HOP -> Pacific Grove PGR
    rep("isla_san_jeronimo",    11), # ISJ
    rep("laguna_beach",         13), # LB
    rep("point_dume",            5), # PD
    rep("santa_barbara",        13), # SB
    rep("la_jolla_san_diego",   13), # SD
    rep("tomales_bay",           1), # TB
    rep("redondo_beach",        10), # RB
    rep("santa_cruz_harbour",   10)  # SCH
  ), c(
    rep("bodega_bay",            5), # BB
    rep("big_creek",            13), # BIGC -> BCR
    rep("elkhorn",               6), # ELK
    rep("pacific_grove",        13), # HOP -> Pacific Grove PGR
    rep("laguna_beach",         13), # LB
    rep("point_dume",            5), # PD
    rep("santa_barbara",        13), # SB
    rep("san_clemente_island",  10), # SCL
    rep("la_jolla_san_diego",   13), # SD
    rep("tomales_bay",           1), # TB
    rep("redondo_beach",        10), # RB
    rep("santa_cruz_harbour",   10)  # SCH
  ), c(
    rep("catalina_island",      11), # CAT
    rep("guadalupe_island",     10)  # GUA
  ), c(
    rep("guadalupe_island",     10), # GUA
    rep("isla_san_jeronimo",    11)  # ISJ
  ), c(
    rep("guadalupe_island",     10), # GUA
    rep("san_clemente_island",  10)  # SCL
  ), c(
    rep("catalina_island",      11), # CAT
    rep("guadalupe_island",     10), # GUA
    rep("isla_san_jeronimo",    11), # ISJ
    rep("san_clemente_island",  10)  # SCL
  ), c(
    rep("bodega_bay",            5), # BB
    rep("big_creek",            13), # BIGC -> BCR
    rep("elkhorn",               6), # ELK
    rep("pacific_grove",        13), # HOP -> Pacific Grove PGR
    rep("santa_barbara",        13), # SB
    rep("tomales_bay",           1), # TB
    rep("santa_cruz_harbour",   10)  # SCH
  ), c(
    rep("isla_san_jeronimo",    11), # ISJ
    rep("san_clemente_island",  10)  # SCL
  ), c(
    rep("laguna_beach",         13), # LB
    rep("point_dume",            5), # PD
    rep("santa_barbara",        13), # SB
    rep("la_jolla_san_diego",   13), # SD
    rep("redondo_beach",        10) # RB
  )
)
```

```{r}
for (i in 1:dim(bed_files)[1]) {
  pcadapt <- here("plink2", bed_files$paths[[i]]) |> 
    read.pcadapt(type = "bed") |> 
    pcadapt(K = 20)
  
  pcadapt |>
    pcadapt_screeplot() |> 
    ggsave(
      here("output", "plots", paste0("pcadapt_", bed_files$suffixes[[i]], "_screeplot.pdf")),
      plot = _,
      width = 10, height = 8
    )
  
  for (j in 1:19) {
    pcadapt |>
      plot(option = "scores", pop = bed_files$popmaps[[i]], i = j, j = j + 1) |> 
      ggsave(
        here("output", "plots", "pcadapt",
             paste0("pcadapt_", bed_files$suffixes[[i]], "_PCA_", as.character(j),"_", as.character(j + 1), ".pdf")),
        plot = _,
        width = 10, height = 8
      )
  }
}

# map2(bed_files$paths, bed_files$suffixes,
#      \(x, y) {
#        here("plink2", x) |> 
#          read.pcadapt(type = "bed") |> 
#          pcadapt(K = 20) |> 
#          \(z) {
#            z |> 
#              pcadapt_screeplot() |> 
#              ggsave(
#                here("output", "plots", paste0("pcadapt_", y, "_screeplot.pdf")),
#                plot = _,
#                width = 10, height = 8
#              )
#            for (i in 1:19) {
#              z |> plot(option = "scores", pop = popmap_lcwgs, i = i, j = i + 1)
#            }
#          }
#          
#      })
```

## All populations

```{r}
#| label: pcadapt

popmap_lcwgs <- c(
  rep("bodega_bay",            5), # BB
  rep("big_creek",            13), # BIGC -> BCR
  rep("catalina_island",      11), # CAT
  rep("elkhorn",               6), # ELK
  rep("guadalupe_island",     10), # GUA
  rep("pacific_grove",        13), # HOP -> Pacific Grove PGR 
  rep("isla_san_jeronimo",    11), # ISJ
  rep("laguna_beach",         13), # LB
  rep("point_dume",            5), # PD
  rep("santa_barbara",        13), # SB
  rep("san_clemente_island",  10), # SCL
  rep("la_jolla_san_diego",   13), # SD
  rep("tomales_bay",           1), # TB
  rep("redondo_beach",        10), # RB
  rep("santa_cruz_harbour",   10) # SCH
)

pcadapt_lcwgs <- here("plink2", "8-Embiotoca_filtered.bed") |> 
  read.pcadapt(type = "bed") |>
  pcadapt(K = 6)
```

```{r}
#| eval: false

write_rds(pcadapt_lcwgs, file = here("output", "pcadapt_lcwgs_K9.rds"))
```

```{r}
#| label: ExploPlots

for (i in 1:19) {
  pcadapt_lcwgs |> plot(option = "scores", pop = popmap_lcwgs, i = i, j = i + 1)
}

pcadapt_lcwgs |> plot(option = "scores", pop = popmap_lcwgs, i = 1, j = 2) |> ggsave(
        here("output", "plots", paste0("pcadapt_all_PCA_1_2.pdf")),
        plot = _,
        width = 10, height = 8
      )
  

pcadapt_lcwgs |> plot(option = "qqplot")

pcadapt_lcwgs$pvalues |> hist(
  xlab   = "p-values",
  main   = NULL,
  breaks = 50,
  col    = "orange"
)
```

```{r}
colnames(pcadapt_lcwgs$loadings) <- paste0("PC", seq(1, 6))

for (i in paste0("PC", seq(1, 6))) {
  ggsave(
    plot = pcadapt_lcwgs$loadings |>
      data.frame() |> 
      mutate(index = row_number()) |> 
      ggplot() +
      geom_point(aes_string(x = "index", y = i)) +
      theme_bw(),
    filename  = here(paste0("pcadapt_all_pops_K9_loadings_", i, ".png")),
    width     = 60,
    limitsize = FALSE
  )
}

```

## Coastal populations

```{r}
#| label: Coastpcadapt

popmap_lcwgs_coast <- c(
  rep("bodega_bay",            5), # BB
  rep("big_creek",            13), # BIGC -> BCR
  rep("elkhorn",               6), # ELK
  rep("pacific_grove",        13), # HOP -> Pacific Grove PGR
  rep("laguna_beach",         13), # LB
  rep("point_dume",            5), # PD
  rep("santa_barbara",        13), # SB
  rep("la_jolla_san_diego",   13), # SD
  rep("tomales_bay",           1), # TB
  rep("redondo_beach",        10), # RB
  rep("santa_cruz_harbour",   10)  # SCH
)

pcadapt_lcwgs_coast <- pcadapt(EJA_lcwgs_coast, K = 5)
```

```{r}
#| eval: false

pcadapt_lcwgs_coast |> 
  write_rds(file = here("output", "pcadapt_lcwgs_coast_K5.rds"))
```

```{r}
#| label: CoastExploPlots

pcadapt_lcwgs_coast |> plot(option = "manhattan")

pcadapt_lcwgs_coast |>
  plot(option = "scores", pop = popmap_lcwgs_coast, i = 1, j = 2)

pcadapt_lcwgs_coast |>
  plot(option = "scores", pop = popmap_lcwgs_coast, i = 2, j = 3)

pcadapt_lcwgs_coast |>
  plot(option = "scores", pop = popmap_lcwgs_coast, i = 3, j = 4)

pcadapt_lcwgs_coast |>
  plot(option = "scores", pop = popmap_lcwgs_coast, i = 4, j = 5)

plot(pcadapt_lcwgs_coast, option = "qqplot") + theme_minimal()

pcadapt_lcwgs_coast$pvalues |> hist(
  xlab   = "p-values",
  main   = NULL,
  breaks = 50,
  col    = "orange"
)
```

```{r}
#| label: CoastMemclean

rm(EJA_lcwgs_coast, coast_PCA_1_2, coast_manhattan)
```

## North Coastal populations

```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

./plink2 --bfile 8-Embiotoca_filtered --keep whitelist_north.txt --make-bed --out 8-Embiotoca_filtered_north
```

```{r}
#| label: Northpcadapt

popmap_lcwgs_north <- c(
  rep("bodega_bay",            5), # BB
  rep("big_creek",            13), # BIGC -> BCR
  rep("elkhorn",               6), # ELK
  rep("pacific_grove",        13), # HOP -> Pacific Grove PGR
  rep("santa_barbara",        13), # SB
  rep("tomales_bay",           1), # TB
  rep("santa_cruz_harbour",   10)  # SCH
)

pcadapt_lcwgs_north <- pcadapt(EJA_lcwgs_north, K = 3)
```

```{r}
#| eval: false

write_rds(pcadapt_lcwgs_north, file = here("output", "pcadapt_lcwgs_north.rds"))
```

```{r}
#| label: NorthExploPlots

north_PCA_1_2 <- pcadapt_lcwgs_north |>
  plot(option = "scores", pop = popmap_lcwgs_north, i = 1, j = 2) +
  theme_minimal()

save_open_plot(
  here("output", "plots", "north_PCA_1_2.pdf"),
  plot   = north_PCA_1_2,
  width  = 10,
  height = 8
)

plot(pcadapt_lcwgs_north, option = "qqplot") + theme_minimal()

pcadapt_lcwgs_north$pvalues |> hist(
  xlab   = "p-values",
  main   = NULL,
  breaks = 50,
  col    = "orange"
)
```

```{r}
#| label: NorthMemclean

rm(EJA_lcwgs_north, north_PCA_1_2, north_manhattan)
```

## South Coastal populations

```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

./plink2 --bfile 8-Embiotoca_filtered --keep whitelist_south.txt --make-bed --out 8-Embiotoca_filtered_south
```

```{r}
#| label: Southpcadapt

popmap_lcwgs_south <- c(
  rep("laguna_beach",         13), # LB
  rep("point_dume",            5), # PD
  rep("santa_barbara",        13), # SB
  rep("la_jolla_san_diego",   13), # SD
  rep("redondo_beach",        10) # RB
)

pcadapt_lcwgs_south <- pcadapt(EJA_lcwgs_south, K = 2)
```

```{r}
#| eval: false

write_rds(pcadapt_lcwgs_south, file = here("output", "pcadapt_lcwgs_south.rds"))
```

```{r}
#| label: SouthExploPlots

south_PCA_1_2 <- pcadapt_lcwgs_south |>
  plot(option = "scores", pop = popmap_lcwgs_south, i = 1, j = 2) +
  theme_minimal()

save_open_plot(
  here("output", "plots", "south_PCA_1_2.pdf"),
  plot   = south_PCA_1_2,
  width  = 10,
  height = 8
)

plot(pcadapt_lcwgs_south, option = "qqplot") + theme_minimal()

pcadapt_lcwgs_south$pvalues |> hist(
  xlab   = "p-values",
  main   = NULL,
  breaks = 50,
  col    = "orange"
)
```

```{r}
#| label: CoastMemclean

rm(EJA_lcwgs_south, south_PCA_1_2, south_manhattan)
```

## Islands

```{r}
#| label: ISLpcadapt

popmap_lcwgs_isl <- c(
  rep("catalina_island",      11), # CAT
  rep("guadalupe_island",     10), # GUA
  rep("isla_san_jeronimo",    11), # ISJ
  rep("san_clemente_island",  10)  # SCL
)

pcadapt_lcwgs_isl <- lcwgs_isl |> pcadapt(K = 3)
```

```{r}
#| eval: false

pcadapt_lcwgs_isl |>
  write_rds(file = here("output", "pcadapt_lcwgs_isl.rds"))
```

```{r}
#| label: ISLExploPlots

isl_PCA_1_2 <- pcadapt_lcwgs_isl |>
  plot(option = "scores", pop = popmap_lcwgs_isl, i = 1, j = 2) +
  theme_minimal()

save_open_plot(
  here("output", "plots", "isl_PCA_1_2.pdf"),
  plot   = isl_PCA_1_2,
  width  = 10,
  height = 8
)

isl_PCA_2_3 <- pcadapt_lcwgs_isl |>
  plot(option = "scores", pop = popmap_lcwgs_isl, i = 2, j = 3) +
  theme_minimal()

save_open_plot(
  here("output", "plots", "isl_PCA_2_3.pdf"),
  plot   = isl_PCA_2_3,
  width  = 10,
  height = 8
)

pcadapt_lcwgs_isl |> plot(option = "qqplot") + theme_minimal()

pcadapt_lcwgs_isl$pvalues |> hist(
  xlab   = "p-values",
  main   = NULL,
  breaks = 50,
  col    = "orange"
)
```

```{r}
#| label: ISLMemclean

rm(lcwgs_isl, isl_PCA_1_2, isl_PCA_2_3, isl_manhattan)
```

## Coast + ISJ

```{r}
#| label: CoastISJpcadapt

popmap_lcwgs_coast_isj <- c(
  rep("bodega_bay",            5), # BB
  rep("big_creek",            13), # BIGC -> BCR
  rep("elkhorn",               6), # ELK
  rep("pacific_grove",        13), # HOP -> Pacific Grove PGR
  rep("isla_san_jeronimo",    11), # ISJ
  rep("laguna_beach",         13), # LB
  rep("point_dume",            5), # PD
  rep("santa_barbara",        13), # SB
  rep("la_jolla_san_diego",   13), # SD
  rep("tomales_bay",           1), # TB
  rep("redondo_beach",        10), # RB
  rep("santa_cruz_harbour",   10)  # SCH
)

pcadapt_lcwgs_coast_isj <- EJA_lcwgs_coast_isj |> pcadapt(K = 6)
```

```{r}
#| eval: false

pcadapt_lcwgs_coast_isj |> 
  write_rds(file = here("output", "pcadapt_lcwgs_coast_isj_K6.rds"))
```

```{r}
#| label: CoastISJExploPlots

pcadapt_lcwgs_coast_isj |>
  plot(option = "scores", pop = popmap_lcwgs_coast_isj, i = 1, j = 2)
pcadapt_lcwgs_coast_isj |>
  plot(option = "scores", pop = popmap_lcwgs_coast_isj, i = 2, j = 3)
pcadapt_lcwgs_coast_isj |>
  plot(option = "scores", pop = popmap_lcwgs_coast_isj, i = 3, j = 4)
pcadapt_lcwgs_coast_isj |>
  plot(option = "scores", pop = popmap_lcwgs_coast_isj, i = 4, j = 5)
pcadapt_lcwgs_coast_isj |>
  plot(option = "scores", pop = popmap_lcwgs_coast_isj, i = 5, j = 6)

pcadapt_lcwgs_coast_isj |> plot(option = "qqplot") + theme_minimal()

pcadapt_lcwgs_coast_isj$pvalues |> hist(
  xlab   = "p-values",
  main   = NULL,
  breaks = 50,
  col    = "orange"
)
```

```{r}
#| label: CoastMemclean

rm(EJA_lcwgs_coast_isj, coast_isj_PCA_1_2, coast_isj_PCA_2_3, coast_isj_manhattan)
```

## Coast + SCL

```{r}
#| label: CoastSCLpcadapt

popmap_lcwgs_coast_scl <- c(
  rep("bodega_bay",            5), # BB
  rep("big_creek",            13), # BIGC -> BCR
  rep("elkhorn",               6), # ELK
  rep("pacific_grove",        13), # HOP -> Pacific Grove PGR
  rep("laguna_beach",         13), # LB
  rep("point_dume",            5), # PD
  rep("santa_barbara",        13), # SB
  rep("san_clemente_island",  10), # SCL
  rep("la_jolla_san_diego",   13), # SD
  rep("tomales_bay",           1), # TB
  rep("redondo_beach",        10), # RB
  rep("santa_cruz_harbour",   10)  # SCH
)

pcadapt_lcwgs_coast_scl <- lcwgs_coast_scl |> pcadapt(K = 6)
```

```{r}
#| eval: false

pcadapt_lcwgs_coast_scl |> 
  write_rds(file = here("output", "pcadapt_lcwgs_coast_scl_K6.rds"))
```

```{r}
#| label: CoastSCLExploPlots

pcadapt_lcwgs_coast_scl |>
  plot(option = "scores", pop = popmap_lcwgs_coast_scl, i = 1, j = 2)

pcadapt_lcwgs_coast_scl |>
  plot(option = "scores", pop = popmap_lcwgs_coast_scl, i = 2, j = 3)

pcadapt_lcwgs_coast_scl |>
  plot(option = "scores", pop = popmap_lcwgs_coast_scl, i = 3, j = 4)

pcadapt_lcwgs_coast_scl |>
  plot(option = "scores", pop = popmap_lcwgs_coast_scl, i = 4, j = 5)

pcadapt_lcwgs_coast_scl |>
  plot(option = "scores", pop = popmap_lcwgs_coast_scl, i = 5, j = 6)

pcadapt_lcwgs_coast_scl |> plot(option = "qqplot") + theme_minimal()

pcadapt_lcwgs_coast_scl$pvalues |> hist(
  xlab   = "p-values",
  main   = NULL,
  breaks = 50,
  col    = "orange"
)
```

```{r}
#| label: CoastSCLMemclean

rm(lcwgs_coast_scl, coast_scl_PCA_1_2, coast_scl_PCA_2_3, coast_scl_manhattan)
```

## Coast + GUA

```{r}
#| label: CoastGUApcadapt

popmap_lcwgs_coast_gua <- c(
  rep("bodega_bay",            5), # BB
  rep("big_creek",            13), # BIGC -> BCR
  rep("elkhorn",               6), # ELK
  rep("guadalupe_island",     10), # GUA
  rep("pacific_grove",        13), # HOP -> Pacific Grove PGR
  rep("laguna_beach",         13), # LB
  rep("point_dume",            5), # PD
  rep("santa_barbara",        13), # SB
  rep("la_jolla_san_diego",   13), # SD
  rep("tomales_bay",           1), # TB
  rep("redondo_beach",        10), # RB
  rep("santa_cruz_harbour",   10)  # SCH
)

pcadapt_lcwgs_coast_gua <- lcwgs_coast_gua |> pcadapt(K = 6)
```

```{r}
#| eval: false

pcadapt_lcwgs_coast_gua |>
  write_rds(file = here("output", "pcadapt_lcwgs_coast_gua_K6.rds"))
```

```{r}
#| label: CoastGUAExploPlots

coast_gua_PCA_1_2 <- pcadapt_lcwgs_coast_gua |>
  plot(option = "scores", pop = popmap_lcwgs_coast_gua, i = 1, j = 2)

coast_gua_PCA_2_3 <- pcadapt_lcwgs_coast_gua |>
  plot(option = "scores", pop = popmap_lcwgs_coast_gua, i = 2, j = 3)

coast_gua_PCA_2_3 <- pcadapt_lcwgs_coast_gua |>
  plot(option = "scores", pop = popmap_lcwgs_coast_gua, i = 3, j = 4)

coast_gua_PCA_2_3 <- pcadapt_lcwgs_coast_gua |>
  plot(option = "scores", pop = popmap_lcwgs_coast_gua, i = 4, j = 5)

coast_gua_PCA_2_3 <- pcadapt_lcwgs_coast_gua |>
  plot(option = "scores", pop = popmap_lcwgs_coast_gua, i = 5, j = 6)

pcadapt_lcwgs_coast_gua |> plot(option = "qqplot") + theme_minimal()

pcadapt_lcwgs_coast_gua$pvalues |> hist(
  xlab   = "p-values",
  main   = NULL,
  breaks = 50,
  col    = "orange"
)
```

```{r}
#| label: CoastGUAMemclean

rm(lcwgs_coast_gua, coast_gua_PCA_1_2, coast_gua_PCA_2_3, coast_gua_manhattan)
```

## Coast + CAT

```{r}
#| label: CoastCATpcadapt

popmap_lcwgs_coast_cat <- c(
  rep("bodega_bay",            5), # BB
  rep("big_creek",            13), # BIGC -> BCR
  rep("catalina_island",      11), # CAT
  rep("elkhorn",               6), # ELK
  rep("pacific_grove",        13), # HOP -> Pacific Grove PGR
  rep("laguna_beach",         13), # LB
  rep("point_dume",            5), # PD
  rep("santa_barbara",        13), # SB
  rep("la_jolla_san_diego",   13), # SD
  rep("tomales_bay",           1), # TB
  rep("redondo_beach",        10), # RB
  rep("santa_cruz_harbour",   10)  # SCH
)

pcadapt_lcwgs_coast_cat <- lcwgs_coast_cat |> pcadapt(K = 6)
```

```{r}
#| eval: false

pcadapt_lcwgs_coast_cat |> 
  write_rds(file = here("output", "pcadapt_lcwgs_coast_cat_K6.rds"))
```

```{r}
#| label: CoastCATExploPlots

pcadapt_lcwgs_coast_cat |>
  plot(option = "scores", pop = popmap_lcwgs_coast_cat, i = 1, j = 2)

pcadapt_lcwgs_coast_cat |>
  plot(option = "scores", pop = popmap_lcwgs_coast_cat, i = 2, j = 3)
  
pcadapt_lcwgs_coast_cat |>
  plot(option = "scores", pop = popmap_lcwgs_coast_cat, i = 3, j = 4)

pcadapt_lcwgs_coast_cat |>
  plot(option = "scores", pop = popmap_lcwgs_coast_cat, i = 4, j = 5)

pcadapt_lcwgs_coast_cat |>
  plot(option = "scores", pop = popmap_lcwgs_coast_cat, i = 5, j = 6)


pcadapt_lcwgs_coast_cat |> plot(option = "qqplot") + theme_minimal()

pcadapt_lcwgs_coast_cat$pvalues |> hist(
  xlab   = "p-values",
  main   = NULL,
  breaks = 50,
  col    = "orange"
)
```

```{r}
#| label: CoastCATMemclean

rm(lcwgs_coast_cat, coast_cat_PCA_1_2, coast_cat_PCA_2_3, coast_cat_manhattan)
```

## CAT + SCL

```{r}
#| label: CATSCLpcadapt

popmap_lcwgs_cat_scl <- c(
  rep("catalina_island",      11), # CAT
  rep("san_clemente_island",  10)  # SCL
)

pcadapt_lcwgs_cat_scl <- lcwgs_cat_scl |> pcadapt(K = 1)
```

```{r}
#| eval: false

pcadapt_lcwgs_cat_scl |> 
  write_rds(file = here("output", "pcadapt_lcwgs_cat_scl.rds"))
```

```{r}
#| label: CATSCLExploPlots

pcadapt_lcwgs_cat_scl |> plot(option = "qqplot") + theme_minimal()

pcadapt_lcwgs_cat_scl$pvalues |> hist(
  xlab   = "p-values",
  main   = NULL,
  breaks = 50,
  col    = "orange"
)
```

```{r}
#| label: CoastCATMemclean

rm(lcwgs_cat_scl, cat_scl_manhattan)
```

## GUA + CAT

```{r}
#| label: GUACATpcadapt

popmap_lcwgs_gua_cat <- c(
  rep("catalina_island",      11), # CAT
  rep("guadalupe_island",     10)  # GUA
)

pcadapt_lcwgs_gua_cat <- lcwgs_gua_cat |> pcadapt(K = 1)
```

```{r}
#| eval: false

pcadapt_lcwgs_gua_cat |> 
  write_rds(file = here("output", "pcadapt_lcwgs_gua_cat.rds"))
```

```{r}
#| label: GUACATExploPlots

pcadapt_lcwgs_gua_cat |> plot(option = "qqplot") + theme_minimal()

pcadapt_lcwgs_gua_cat$pvalues |> hist(
  xlab   = "p-values",
  main   = NULL,
  breaks = 50,
  col    = "orange"
)
```

```{r}
#| label: GUACATMemclean

rm(lcwgs_gua_cat, gua_cat_manhattan)
```

## GUA + SCL

```{r}
#| label: GUASCLpcadapt

popmap_lcwgs_gua_scl <- c(
  rep("guadalupe_island",     10), # GUA
  rep("san_clemente_island",  10)  # SCL
)

pcadapt_lcwgs_gua_scl <- lcwgs_gua_scl |> pcadapt(K = 1)
```

```{r}
#| eval: false

pcadapt_lcwgs_gua_scl |> 
  write_rds(file = here("output", "pcadapt_lcwgs_gua_scl.rds"))
```

```{r}
#| label: GUASCLExploPlots

pcadapt_lcwgs_gua_scl |> plot(option = "qqplot") + theme_minimal()

pcadapt_lcwgs_gua_scl$pvalues |> hist(
  xlab   = "p-values",
  main   = NULL,
  breaks = 50,
  col    = "orange"
)
```

```{r}
#| label: GUASCLMemclean

rm(lcwgs_gua_scl, gua_scl_manhattan)
```

## GUA + ISJ

```{r}
#| label: GUAISJpcadapt

popmap_lcwgs_GUA_ISJ <- c(
  rep("guadalupe_island",     10), # GUA
  rep("isla_san_jeronimo",    11)  # ISJ
)

pcadapt_lcwgs_GUA_ISJ <- lcwgs_gua_isj |> pcadapt(K = 1)
```

```{r}
#| eval: false

pcadapt_lcwgs_GUA_ISJ |> 
  write_rds(file = here("output", "pcadapt_lcwgs_gua_isj.rds"))
```

```{r}
#| label: GUAISJExploPlots

pcadapt_lcwgs_GUA_ISJ |> plot(option = "qqplot") + theme_minimal()

pcadapt_lcwgs_GUA_ISJ$pvalues |> hist(
  xlab   = "p-values",
  main   = NULL,
  breaks = 50,
  col    = "orange"
)
```

```{r}
#| label: GUAISJMemclean

rm(lcwgs_GUA_ISJ, GUA_ISJ_manhattan)
```

## CAT + ISJ

```{r}
#| label: CATISJpcadapt

popmap_lcwgs_cat_isj <-  c(
  rep("catalina_island",      11), # CAT
  rep("isla_san_jeronimo",    11)  # ISJ
)

pcadapt_lcwgs_cat_isj <- lcwgs_cat_isj |> pcadapt(K = 1)
```

```{r}
#| eval: false

pcadapt_lcwgs_cat_isj |> 
  write_rds(file = here("output", "pcadapt_lcwgs_cat_isj.rds"))
```

```{r}
#| label: CATISJExploPlots

pcadapt_lcwgs_cat_isj |> plot(option = "qqplot") + theme_minimal()

pcadapt_lcwgs_cat_isj$pvalues |> hist(
  xlab   = "p-values",
  main   = NULL,
  breaks = 50,
  col    = "orange"
)
```

```{r}
#| label: CATISJMemclean

rm(lcwgs_cat_isj, cat_isj_manhattan)
```

## SCL + ISJ

```{r}
#| label: SCLISJpcadapt

popmap_lcwgs_scl_isj <- c(
  rep("isla_san_jeronimo",    11), # ISJ
  rep("san_clemente_island",  10)  # SCL
)

pcadapt_lcwgs_scl_isj <- lcwgs_scl_isj |> pcadapt(K = 1)
```

```{r}
#| eval: false

pcadapt_lcwgs_scl_isj |> 
  write_rds(file = here("output", "pcadapt_lcwgs_scl_isj.rds"))
```

```{r}
#| label: SCLISJExploPlots

pcadapt_lcwgs_scl_isj |> plot(option = "qqplot") + theme_minimal()

pcadapt_lcwgs_scl_isj$pvalues |> hist(
  xlab   = "p-values",
  main   = NULL,
  breaks = 50,
  col    = "orange"
)
```

```{r}
#| label: SCLISJMemclean

rm(lcwgs_scl_isj, scl_isj_manhattan)
```

# Outliers determination and annotation

## Load species Metadata

```{r}
EJA_annotations <- here("data", "EJA_annotations.gff") |> # Load annotated genome
  read.gff() |>
  as_tibble() |> 
  filter(type != "region")

chromosomes_info <- here("data", "JAKOON01_catalog", "sequence_report.tsv") |>  # Load misc inform
  read_tsv() |>
  select(c(`GenBank seq accession`, `RefSeq seq accession`))
```

### All outliers

```{r}
read_write_locations <- tibble(
  read_files = list.files(here("output")) |> str_subset("pcadapt")
) |> 
  mutate(
    write_files = read_files |>
      str_replace("pcadapt", "EJA") |>
      str_replace(".rds", "_1E-2_outliers_annotations.csv")
  )

map2(
  read_write_locations$read_files, read_write_locations$write_files,
     \(x, y) {
       print(x)
       
       here("output", x) |>
         readRDS() |>
         Get_outliers(method = "bonferroni", pvalue_threshold = .01, minimize_length = FALSE) |> 
         Get_annotations(chrom_info = chromosomes_info, gff_annotation = EJA_annotations) |>
         Clean_annotations() |> 
         write_csv2(here("output", y))
     }
)
```

Summary is saved in `EJA_lcwgs_outliers_log.log`.

## For only one

```{r}
test <- here("output", "pcadapt_lcwgs_K9.rds") |>
  readRDS() |>
  Get_outliers(method = "bonferroni", pvalue_threshold = .001, minimize_length = FALSE) |>
  Get_annotations(chrom_info = chromosomes_info, gff_annotation = EJA_annotations) |>
  Clean_annotations() |> 
  write_csv2(here("output", "EJA_lcwgs_cat_isj_5E-2_outliers_annotations.csv"))
```

## Add Principal Components association to tables

```{r}
tibble(
  read_csv     = list.files(here("output")) |> str_subset("outliers_annotations"),
  read_pcadapt = list.files(here("output")) |> str_subset("pcadapt")
) |> 
  (\(x) {
    map2(
      x$read_csv, x$read_pcadapt,
      \(x, y) {
        here("output", x) |> 
          read_csv2() |> 
          inner_join(
            readRDS(file = here("data", "SNP_positions.rds")),
            by = join_by("GenBank seq accession" == "CHROM", "POS")
          ) |> 
          mutate(
            PC = here("output", y) |> read_rds() |> get.pc(list = rank) |> pull(PC)
          ) |> 
          select(-c(rank)) |> 
          write_csv2(file = here("output", sub(".csv", "_PC.csv", x)))
      }
    )
  })()
```

## Add Allele frequencies

### Send-away code

```{r}
#| eval: false

EJA_gen <- readRDS(here("data", "EJA_lcwgs_gen.rds"))

EJA_allele_frequencies <- makefreq(EJA_gen)

EJA_allele_frequencies_tb <- EJA_allele_frequencies |>
    t() |>
    as_tibble(rownames = "POSITION") |>
    separate_wider_delim(POSITION, delim = "_", names = c("CHROM", "TBD", "POSALLELE")) |>
    separate_wider_delim(POSALLELE, delim = ".", names = c("POS", "ALLELE")) |>
    select(-c(TBD)) |>
    mutate(CHROM = paste0(CHROM, ".1"))

saveRDS(EJA_allele_frequencies_tb, here("output", "EJA_allele_frequencies_tb.rds"))
```

### Add results to tables

```{r}
EJA_allele_freq <- readRDS(here("output", "EJA_allele_frequencies_tb.rds")) |> 
  filter(ALLELE == 0) |> 
  select(-ALLELE)

list.files(here("output")) |>
  str_subset("outliers_annotations_PC") |> 
  map(\(x) {
    here("output", x) |> 
      read_csv2() |> 
      left_join(
        EJA_allele_freq,
        by = join_by("GenBank seq accession" == "CHROM", "POS")
      ) |> 
      write_csv2(file = here("output", sub(".csv", "_allele_freq.csv", x)))
  })
```


### Get outliers fasta
#### Read genome
```{r}
lcwgs_catalog <- read.dna(
    here("data", "JAKOON01_catalog", "catalog.fa"),
    format       = "fasta",
    as.character = TRUE
  )

names(lcwgs_catalog) <- names(lcwgs_catalog) |>
  str_split(pattern = " ") |>
  map(\(x) x[1]) |>
  unlist()
```

#### Get corresponding read windows
```{r}
chromosomes_length <- read_tsv(here("data", "JAKOON01_catalog", "JAKOON01_contigs.tsv"))

pvalues_positions_filtered_dup_sequences <- pvalues_positions_filtered_dup |>
  mutate(
    sequences = row_number() |>
      map(
        \(x) lcwgs_catalog[[CHROM[[x]]]][Get_window(CHROM[[x]], POS[[x]])] |>
          paste(collapse = "") |>
          toupper()
      ) |> 
      unlist()
    )
```

#### Write to fasta files
```{r}
pvalues_positions_filtered_dup_sequences |>
  mutate(seq_name = paste0(">", CHROM, " POS: ", as.character(POS))) |>
  select(seq_name, sequences) |> 
  filter(row_number() %in% 1:200) |> 
  write.table(
    file      = here("output", "lcwgs_outliers_sequences_1.fa"),
    sep       = "\n",
    row.names = FALSE,
    col.names = FALSE,
    quote     = FALSE
  )

pvalues_positions_filtered_dup_sequences |>
  mutate(seq_name = paste0(">", CHROM, " POS: ", as.character(POS))) |>
  select(seq_name, sequences) |> 
  filter(row_number() %in% 201:400) |> 
  write.table(
    file      = here("output", "lcwgs_outliers_sequences_2.fa"),
    sep       = "\n",
    row.names = FALSE,
    col.names = FALSE,
    quote     = FALSE
  )

pvalues_positions_filtered_dup_sequences |>
  mutate(seq_name = paste0(">", CHROM, " POS: ", as.character(POS))) |>
  select(seq_name, sequences) |> 
  filter(row_number() %in% 401:456) |> 
  write.table(
    file      = here("output", "lcwgs_outliers_sequences_3.fa"),
    sep       = "\n",
    row.names = FALSE,
    col.names = FALSE,
    quote     = FALSE
  )
```

```{r}
rm(chromosomes_length,
   pvalues_positions_filtered_dup_sequences)
gc()
```

# Outliers Protein sequences for KEGG

### Get BLAST results

```{r}
pvalues_positions <- pvalues_positions |>
  mutate(Rank = paste0(CHROM, " POS: ", as.character(POS)))

lcwgs_EJA_outliers_blast_tib <- c("GZSEDPGT016-Alignment.xml",
                                  "GZSJHA6W016-Alignment.xml",
                                  "GZSMYB0J016-Alignment.xml") |> # "Single-file XML2" file from NCBI BLASTn output is expected
  here("data", "BLASTs") |>
  map( \(x) read_xml(x) |> XML_to_df()) |> 
  bind_rows() |>
  left_join(pvalues_positions)

rm(pvalues_positions)
```

#### Filter BLAST results

```{r}
lcwgs_EJA_outliers_blast_tib_filtered <- lcwgs_EJA_outliers_blast_tib |>
  filter(bit_score >= 80) |> 
  filter(!grepl("genome assembly", title)) |> 
  filter(!duplicated(Rank)) |> # Keep the first match for each locus
  filter(!duplicated(accession_number))
```

```{r}
write_csv2(
  lcwgs_EJA_outliers_blast_tib_filtered,
  file = here("output", "lcwgs_outliers_match.csv")
)
```

#### Get protein sequences from GenBank

```{r}
lcwgs_protein_sequences <- lcwgs_EJA_outliers_blast_tib_filtered |> Get_protein_sequences()
```

#### Save output as fasta file which can be sent to [blast KOALA](https://www.kegg.jp/blastkoala/).

```{r}
lcwgs_protein_sequences |> 
  mutate(accession_number = paste0("> ", accession_number)) |>
  write.table(
    file      = here("output", "lwcgs_EJA_outliers_protein_sequences.fa"),
    sep       = "\n",
    row.names = FALSE,
    col.names = FALSE,
    quote     = FALSE
  )
```

### Get BLAST results A. ocellaris

```{r}
pvalues_positions <- readRDS(
  file = here("data", "SNP_positions.rds")
) |> 
  mutate(
    pvalues = pcadapt_lcwgs_coast$pvalues,
    logpvalues = Get_pvalues(pcadapt_lcwgs_coast)
  ) |>
  mutate(Rank = paste0(CHROM, " POS: ", as.character(POS)))
```

```{r}
lcwgs_EJA_outliers_blast_A_ocellaris_tib <- c("HFHS4UE9013-Alignment.xml",
                                              "HFHRYX38013-Alignment.xml",
                                              "HFHSG8SF016-Alignment.xml") |> # "Single-file XML2" file from NCBI BLASTn output is expected
  here("data", "BLASTs") |>
  map( \(x) read_xml(x) |> XML_to_df()) |> 
  bind_rows() |>
  left_join(pvalues_positions)

rm(pvalues_positions)
```

```{r}
lcwgs_EJA_outliers_blast_A_ocellaris_tib_filtered <- lcwgs_EJA_outliers_blast_A_ocellaris_tib |>
  filter(bit_score >= 80) |> 
  filter(!grepl("genome assembly", title)) |> 
  filter(!duplicated(Rank)) |> # Keep the first match for each locus
  filter(!duplicated(accession_number))

lcwgs_EJA_outliers_blast_A_ocellaris_tib_filtered |>
  write.csv2(here("output", "lcwgs_outliers_match_A_ocellaris.csv"))
```

# Outliers with `outFLANK`

## Convert all beds to vcfs

```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

bed_names=();
for f in *.bed; do
    bed_names+=("${f%.bed}");
done;

for bed in ${bed_names[@]}; do ./plink2 --bfile $bed --keep-allele-order --recode vcf --out $bed; done;
```

Coast *vs.* Gua and Cat
```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

./plink2 --bfile 8-Embiotoca_filtered --keep whitelist_coastal_gua_cat.txt --make-bed --out 8-Embiotoca_filtered_coastal_gua_cat

./plink2 --bfile 8-Embiotoca_filtered_coastal_gua_cat --keep-allele-order --recode vcf --out 8-Embiotoca_filtered_coastal_gua_cat
```

Coast *vs.* Gua and Scl
```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

./plink2 --bfile 8-Embiotoca_filtered --keep whitelist_coastal_gua_scl.txt --make-bed --out 8-Embiotoca_filtered_coastal_gua_scl

./plink2 --bfile 8-Embiotoca_filtered_coastal_gua_scl --keep-allele-order --recode vcf --out 8-Embiotoca_filtered_coastal_gua_scl
```

Coast *vs.* Gua, Cat and Scl
```{bash}
cd C:/Users/aoules/Desktop/R_projects/Surfperch_GenPop/plink2

./plink2 --bfile 8-Embiotoca_filtered --keep whitelist_coastal_gua_cat_scl.txt --make-bed --out 8-Embiotoca_filtered_coastal_gua_cat_scl

./plink2 --bfile 8-Embiotoca_filtered_coastal_gua_cat_scl --keep-allele-order --recode vcf --out 8-Embiotoca_filtered_coastal_gua_cat_scl
```

## prep

```{r}
library(OutFLANK)
```

```{r}
popmaps <- list(
  "vcf_paths" = list.files(here("plink2"), pattern = ".vcf$"),
  "names"     = list("all", "cat_isj", "cat_scl", "coastal", "coastal_cat",
                     "coastal_gua", "coastal_gua_cat", "coastal_gua_cat_scl",
                     "coastal_gua_scl", "coastal_isj", "coastal_scl", "gua_cat",
                     "gua_isj", "gua_scl", "isl", "north", "scl_isj", "south"),
  "maps"      = list(
    "all" = c(
      rep("bodega_bay",            5), # BB
      rep("big_creek",            13), # BIGC -> BCR
      rep("catalina_island",      11), # CAT
      rep("elkhorn",               6), # ELK
      rep("guadalupe_island",     10), # GUA
      rep("pacific_grove",        13), # HOP -> Pacific Grove PGR 
      rep("isla_san_jeronimo",    11), # ISJ
      rep("laguna_beach",         13), # LB
      rep("point_dume",            5), # PD
      rep("santa_barbara",        13), # SB
      rep("san_clemente_island",  10), # SCL
      rep("la_jolla_san_diego",   13), # SD
      rep("tomales_bay",           1), # TB
      rep("redondo_beach",        10), # RB
      rep("santa_cruz_harbour",   10) # SCH
    ), "cat_isj" = c(
      rep("catalina_island",      11), # CAT
      rep("isla_san_jeronimo",    11)  # ISJ
    ), "cat_scl" = c(
      rep("catalina_island",      11), # CAT
      rep("san_clemente_island",  10)  # SCL
    ), "coastal" = c(
      rep("bodega_bay",            5), # BB
      rep("big_creek",            13), # BIGC -> BCR
      rep("elkhorn",               6), # ELK
      rep("pacific_grove",        13), # HOP -> Pacific Grove PGR
      rep("laguna_beach",         13), # LB
      rep("point_dume",            5), # PD
      rep("santa_barbara",        13), # SB
      rep("la_jolla_san_diego",   13), # SD
      rep("tomales_bay",           1), # TB
      rep("redondo_beach",        10), # RB
      rep("santa_cruz_harbour",   10)  # SCH
    ), "coastal_cat" = c(
      rep("bodega_bay",            5), # BB
      rep("big_creek",            13), # BIGC -> BCR
      rep("catalina_island",      11), # CAT
      rep("elkhorn",               6), # ELK
      rep("pacific_grove",        13), # HOP -> Pacific Grove PGR
      rep("laguna_beach",         13), # LB
      rep("point_dume",            5), # PD
      rep("santa_barbara",        13), # SB
      rep("la_jolla_san_diego",   13), # SD
      rep("tomales_bay",           1), # TB
      rep("redondo_beach",        10), # RB
      rep("santa_cruz_harbour",   10)  # SCH
    ), "coastal_gua" = c(
      rep("bodega_bay",            5), # BB
      rep("big_creek",            13), # BIGC -> BCR
      rep("elkhorn",               6), # ELK
      rep("guadalupe_island",     10), # GUA
      rep("pacific_grove",        13), # HOP -> Pacific Grove PGR
      rep("laguna_beach",         13), # LB
      rep("point_dume",            5), # PD
      rep("santa_barbara",        13), # SB
      rep("la_jolla_san_diego",   13), # SD
      rep("tomales_bay",           1), # TB
      rep("redondo_beach",        10), # RB
      rep("santa_cruz_harbour",   10)  # SCH
    ), "coastal_gua_cat" = c(
      rep("bodega_bay",            5), # BB
      rep("big_creek",            13), # BIGC -> BCR
      rep("catalina_island",      11), # CAT
      rep("elkhorn",               6), # ELK
      rep("guadalupe_island",     10), # GUA
      rep("pacific_grove",        13), # HOP -> Pacific Grove PGR
      rep("laguna_beach",         13), # LB
      rep("point_dume",            5), # PD
      rep("santa_barbara",        13), # SB
      rep("la_jolla_san_diego",   13), # SD
      rep("tomales_bay",           1), # TB
      rep("redondo_beach",        10), # RB
      rep("santa_cruz_harbour",   10)  # SCH
    ), "coastal_gua_cat_scl" = c(
      rep("bodega_bay",            5), # BB
      rep("big_creek",            13), # BIGC -> BCR
      rep("catalina_island",      11), # CAT
      rep("elkhorn",               6), # ELK
      rep("guadalupe_island",     10), # GUA
      rep("pacific_grove",        13), # HOP -> Pacific Grove PGR
      rep("laguna_beach",         13), # LB
      rep("point_dume",            5), # PD
      rep("santa_barbara",        13), # SB
      rep("san_clemente_island",  10), # SCL
      rep("la_jolla_san_diego",   13), # SD
      rep("tomales_bay",           1), # TB
      rep("redondo_beach",        10), # RB
      rep("santa_cruz_harbour",   10)  # SCH
    ), "coastal_gua_scl" = c(
      rep("bodega_bay",            5), # BB
      rep("big_creek",            13), # BIGC -> BCR
      rep("elkhorn",               6), # ELK
      rep("guadalupe_island",     10), # GUA
      rep("pacific_grove",        13), # HOP -> Pacific Grove PGR
      rep("laguna_beach",         13), # LB
      rep("point_dume",            5), # PD
      rep("santa_barbara",        13), # SB
      rep("san_clemente_island",  10), # SCL
      rep("la_jolla_san_diego",   13), # SD
      rep("tomales_bay",           1), # TB
      rep("redondo_beach",        10), # RB
      rep("santa_cruz_harbour",   10)  # SCH
    ), "coastal_isj" = c(
      rep("bodega_bay",            5), # BB
      rep("big_creek",            13), # BIGC -> BCR
      rep("elkhorn",               6), # ELK
      rep("pacific_grove",        13), # HOP -> Pacific Grove PGR
      rep("isla_san_jeronimo",    11), # ISJ
      rep("laguna_beach",         13), # LB
      rep("point_dume",            5), # PD
      rep("santa_barbara",        13), # SB
      rep("la_jolla_san_diego",   13), # SD
      rep("tomales_bay",           1), # TB
      rep("redondo_beach",        10), # RB
      rep("santa_cruz_harbour",   10)  # SCH
    ), "coastal_scl" = c(
      rep("bodega_bay",            5), # BB
      rep("big_creek",            13), # BIGC -> BCR
      rep("elkhorn",               6), # ELK
      rep("pacific_grove",        13), # HOP -> Pacific Grove PGR
      rep("laguna_beach",         13), # LB
      rep("point_dume",            5), # PD
      rep("santa_barbara",        13), # SB
      rep("san_clemente_island",  10), # SCL
      rep("la_jolla_san_diego",   13), # SD
      rep("tomales_bay",           1), # TB
      rep("redondo_beach",        10), # RB
      rep("santa_cruz_harbour",   10)  # SCH
    ), "gua_cat" = c(
      rep("catalina_island",      11), # CAT
      rep("guadalupe_island",     10)  # GUA
    ), "gua_isj" = c(
      rep("guadalupe_island",     10), # GUA
      rep("isla_san_jeronimo",    11)  # ISJ
    ), "gua_scl" =  c(
      rep("guadalupe_island",     10), # GUA
      rep("san_clemente_island",  10)  # SCL
    ), "isl" =  c(
      rep("catalina_island",      11), # CAT
      rep("guadalupe_island",     10), # GUA
      rep("isla_san_jeronimo",    11), # ISJ
      rep("san_clemente_island",  10)  # SCL
    ), "north" = c(
      rep("bodega_bay",            5), # BB
      rep("big_creek",            13), # BIGC -> BCR
      rep("elkhorn",               6), # ELK
      rep("pacific_grove",        13), # HOP -> Pacific Grove PGR
      rep("santa_barbara",        13), # SB
      rep("tomales_bay",           1), # TB
      rep("santa_cruz_harbour",   10)  # SCH
    ), "scl_isj" = c(
      rep("isla_san_jeronimo",    11), # ISJ
      rep("san_clemente_island",  10)  # SCL
    ), "south" =   c(
      rep("laguna_beach",         13), # LB
      rep("point_dume",            5), # PD
      rep("santa_barbara",        13), # SB
      rep("la_jolla_san_diego",   13), # SD
      rep("redondo_beach",        10) # RB
    )
  )
)
```

## Load data and convert to format:

```{r}
#| echo: false

Vectorize(convert_to_outflank)(
  vcf_path  = here("plink2", popmaps$vcf_paths),
  popmap    = popmaps$maps,
  path_save = here(
    "data",
    paste0("8-Embiotoca_filtered_", popmaps$names,"_outflank.rds")
  )
)
```

## Compute Fst and He

```{r}
for (file in list.files(here("data"), pattern = "_outflank.rds$")) {
  if (
    !file.exists(
      here(
          "data",
          str_replace(file, pattern = "outflank.rds", replacement = "HeFst.rds")
      )
    )
  ) {
    print(paste0("Processing ", file))
    
    tryCatch({
      here("data", file) |> 
        readRDS() |> 
        (\(x) MakeDiploidFSTMat(
          t(x$G),
          locusNames = x$position,
          popNames   = x$pop
        ))() |> 
        saveRDS(
          here(
            "data",
            str_replace(file, pattern = "outflank.rds", replacement = "HeFst.rds")
          )
        )
    }, error = function(e){cat("ERROR :",conditionMessage(e), "\n")})

  }
}
```

## Plot FstHe

```{r}
for (file in list.files(here("data"), pattern = "HeFst.rds$")) {
  if (!file.exists(here("output", "plots",
                    paste0("Fst_He_",
                           file |>
                             str_remove(pattern = "8-Embiotoca_filtered_") |>
                             str_remove(pattern = "_HeFst.rds"),
                           ".png")))) {
    print(paste0("Plotting ", file))
    
    ggsave(
      plot = ggplot(readRDS(here("data", file))) +
        geom_point(aes(x = He, y = FST), alpha = .01) +
        labs(x = "Heterozygosity", y = "Fst") +
        theme_classic(),
      filename = here("output", "plots",
                      paste0("Fst_He_",
                             file |>
                               str_remove(pattern = "8-Embiotoca_filtered_") |>
                               str_remove(pattern = "_HeFst.rds"),
                             ".png")),
      width  = 8,
      height = 4
    )
  }
}
```

```{r}
plot(EJA_Fst$FST, EJA_Fst$FSTNoCorr, col = alpha("black", alpha = .01))
abline(0, 1)
```

## Coast vs GUA vs CAT

```{r}
 here("data", "8-Embiotoca_filtered_coastal_gua_cat_outflank.rds") |> 
        readRDS() |> 
        (\(x) MakeDiploidFSTMat(
          t(x$G),
          locusNames = x$position,
          popNames   = x$pop
        ))() |> 
        saveRDS(
          here(
            "data",
            str_replace("8-Embiotoca_filtered_coastal_gua_cat_outflank.rds", pattern = "outflank.rds", replacement = "HeFst.rds")
          )
        )
```

```{r}
 here("data", "8-Embiotoca_filtered_coastal_gua_scl_outflank.rds") |> 
        readRDS() |> 
        (\(x) MakeDiploidFSTMat(
          t(x$G),
          locusNames = x$position,
          popNames   = x$pop
        ))() |> 
        saveRDS(
          here(
            "data",
            str_replace("8-Embiotoca_filtered_coastal_gua_scl_outflank.rds", pattern = "outflank.rds", replacement = "HeFst.rds")
          )
        )
```

```{r}
 here("data", "8-Embiotoca_filtered_coastal_gua_cat_scl_outflank.rds") |> 
        readRDS() |> 
        (\(x) MakeDiploidFSTMat(
          t(x$G),
          locusNames = x$position,
          popNames   = x$pop
        ))() |> 
        saveRDS(
          here(
            "data",
            str_replace("8-Embiotoca_filtered_coastal_gua_cat_scl_outflank.rds", pattern = "outflank.rds", replacement = "HeFst.rds")
          )
        )
```
# Outliers with `bayscan`

```{r}
library(hierfstat)
library(radiator)

# read_csv2(here("data", "individuals_coordinates.csv")) |> 
#   select(c(Indiv, Localisation)) |> 
#   rename(INDIVIDUALS = Indiv, STRATA = Localisation) |> 
#   write_tsv(here("data", "population_map_bsc.txt"))

EJA_geste <- genomic_converter(
  data     = here("data", "8-Embiotoca_filtered.vcf"),
  strata   = here("data", "population_map_bsc.txt"),
  output   = c("bayescan"),
  filename = "8-Embiotoca_filtered_outflank.geste",
  parallel.core = 1
)
```

